\documentclass{ximera}

\input{../preamble.tex}

\author{Bart Snapp}

\title{A topological approach}

\begin{document}
\begin{abstract}
  We define the dimension of a ring based on its topological
  properties.  Sources and references: \cite{sD2008}.
\end{abstract}
\maketitle

\section{Dimension of a topological space}

Given a topological space, we can define its dimension as follows:


\begin{definition}\index{dimension!topological space}
  Let $X$ be a topological space, the \textbf{dimension} of $X$,
  denoted by $\dim(X)$ is defined as
  \[
  \dim(X) :=\sup\left\{d:
  \begin{minipage}{43ex}
    there exists a chain $X_0\supsetneq \dots \supsetneq X_d$ of
    length $d$ of irreducible closed subsets of $X$,
  \end{minipage}
  \right\}.
  \]
\end{definition}


\begin{definition}\index{Noetherian!topological space}
  If $X$ is a topological space, $X$ is called \textbf{Noetherian} if
  any of the following equivalent conditions hold:
  \begin{enumerate}
  \item Every nonempty family of open sets has a maximal element.
  \item Any increasing chain of open sets stabilizes.
  \item Every nonempty family of closed sets has a minimal element.
  \item Any decreasing chain of closed sets stabilizes.
  \end{enumerate}
\end{definition}



%% \begin{proposition}
%%   If $X$ is a Noetherian topological space,
%%   \[
%%   \dim(X) = \sup\dim(X_i)
%%   \]
%%   where $X_i$ is a component of $X$. 
%% \end{proposition}

\section{Krull dimension of a ring}


\begin{proposition}\index{Noetherian!topological space}\index{Noetherian!ring}
  If $R$ is a Noetherian ring, then $\spec(R)$ is a Noetherian
  topological space.
  \begin{proof}\index{Primary Decomposition Theorem}
    Suppose $R$ is Noetherian, then by primary decomposition of
    ideals, Theorem~\ref{T:PDI}, we have
    \[
    (0) = \q_1\cap \dots \cap \q_n
    \]
    where for each $\q_i$, $\ass(R/\q_i) = \{\p_i\}$.  Then we have
    \[
    \spec(R) = V(0) = V(\p_1) \cup\dots \cup V(\p_n). 
    \]
    Since the $V(\p_i)$ are the irreducible components of $\spec(R)$,
    any decreasing chain of closed sets stabilizes.
  \end{proof}
\end{proposition}



\begin{warning}\index{Noetherian!topological space}\index{Noetherian!ring}
  The converse of the above proposition is \textbf{not} true. We leave
  it as an exercise to show that if $k$ is a field, then
  \[
  R = \frac{k[X_1,\dots,X_n,\dots]}{(X_1^2,\dots,X_n^2,\dots)}
  \]
  is not a Noetherian ring but $\spec(R)$ is a Noetherian topological space.
\end{warning}


\begin{definition}[First Notion of Dimension]\index{dimA@$\dim(A)$}\index{dimM@$\dim(M)$}\index{dimension}\index{Krull dimension}
  If $R$ is a ring, the \textbf{Krull dimension}, denoted by
  $\dim(R)$, is the dimension of the topological space $\spec(R)$. To
  be explicit:
  \[
  \dim(R) = \sup \{ d : \text{there exists $\p_0\subsetneq \p_1
    \subsetneq \dots \subsetneq \p_d$ such that $\p_i \in \spec(R)$} \}.
  \]
  This notion of dimension is often simply referred to as the
  \textbf{dimension} of a ring. If $M$ is an $R$-module, then
  \[
  \dim(M) := \dim(R/\ann(M)).
  \]
\end{definition}


\begin{exercise}\index{support}\index{supp@$\supp(M)$}
  Given a ring $R$ and an $R$-module $M$, show that the dimension of
  $M$ is the dimension of the topological space $\supp(M)\subset
  \spec(R)$.
\end{exercise}

\begin{example}
  While
  \[
  \dim(R) = \sup \{ d : \text{there exists $\p_0\subsetneq \p_1 \subsetneq \dots \subsetneq \p_d$ such that $\p_i \in \spec(R)$} \}.
  \]
  Nagata gives an example of a ring $R$ such that $R$ is Noetherian
  but with infinite dimension in \cite{mN1962}.
\end{example}


\begin{example}
  In the ring $\C[X_1,\dots,X_n]$ we have the chain of prime ideals
  \[
  (0) \subsetneq (X_1) \subsetneq \cdots \subsetneq (X_1,\dots,X_n).
  \]
  Thus the dimension of $\C[X_1,\dots,X_n]$ is at least $n$.
\end{example}


Suppose that a ring $R$ has finite dimension. If $\m_1, \dots, \m_r$
are maximal prime ideals in $\spec(R)$, then $\dim(R) = d$ implies
that there exists a chain of length $d$ ending at one of the $\m_i$'s.
Thus for some $i$,
\[
\dim(R_{\m_i}) = \dim(R). 
\]
Hence we see that some questions about the dimension of a ring can be
reduced to questions about local rings.


\begin{remark}
  We can characterize dimension in two useful ways:
  \begin{enumerate}
  \item $\dim(R) = \sup \dim(R_\m)$ where $\m \in \maxspec(R)$.
  \item $\dim(R) = \sup \dim(R/\p)$ where $\p$ is a minimal prime ideal of $R$. 
  \end{enumerate}
\end{remark}


\begin{question}
  Is the dimension of a Noetherian ring finite?
\end{question}

For every question we cannot answer, there is an easier question that we cannot answer.

\begin{question}
  Is the dimension of a local ring finite?
\end{question}

Work on, and we will resolve these questions.


\end{document}
