\documentclass{ximera}

\input{../preamble.tex}

\title{Gr\"obner bases}

\author{Bart Snapp}

\begin{document}
\begin{abstract}
  We give an algorithm for deciding when a polynomial is a member of
  an ideal.  Sources and references: \cite{CLO2007,hS2003}.
\end{abstract}
\maketitle



\section{The ideal membership problem}

Given a ring and ideal in the ring, we would like to know when an
element is a member of the ideal. We'll need some notation.


\begin{definition}
  Consider a polynomial $f\in k[x_1,\dots, x_n]$ with a monomial order
  $>$.
  \begin{itemize}
    \item The \dfn{multidegree} of $f$ is
    \[
    \multideg(f) := \{\valpha: \text{$c_\valpha \x^\valpha$ is the largest term w.r.t.\ $>$} \}.
    \]
  \item The \dfn{lead coefficient} of $f$ is $\lc(f) := c_{\multideg(f)}$.
  \item The \dfn{lead monomial} of $f$ is $\lm(f) := \x^{\multideg(f)}$.
  \item The \dfn{lead term} of $f$ is $\lt(f) := \lc(f)\cdot \lm(f)$.
  \item The \dfn{ideal of lead terms} of an ideal $I\subset k[\x]$ is
    the ideal
    \[
    \lt(I) = (\lt(f): f\in I)\subset k[\x].
    \]
  \end{itemize}
\end{definition}

We have
\begin{macaulay2}
R = ZZ/101[x,y,z];
f = 3*z^2*y^3 + 7*x^2*y;
degree f
leadCoefficient f
leadMonomial f
leadTerm f

R = ZZ/101[x,y,z,MonomialOrder=>Lex];
f = 3*z^2*y^3 + 7*x^2*y;
degree f
leadCoefficient f
leadMonomial f
leadTerm f
\end{macaulay2}


Now suppose we are working in the ring $k[x_1,\dots, x_n]$ and we have an ideal
\[
I = (g_1,\dots, g_m) \subset k[\x].
\]
Here is a basic algorithm that attempts to solve the ideal membership
problem.

\begin{algorithm}[Division algorithm]
  \hfill
  \begin{algorithmic}[1]
    \Procedure{Division}{$f,\{g_1,\dots,g_m\}$} 
    \State $\texttt{quotient} = 0$
    \State $\texttt{remainder} = 0$
    \Repeat
    \If{$\lt(g_i) \cdot h_i = \lt(f)$}
    \State $\texttt{quotient} \from \texttt{quotient} + h_i\cdot g_i$
    \State $f \from f - h_i\cdot g_i$
    \Else
    \State $\texttt{remainder} \from \texttt{remainder} + \lt(f)$
    \State $f\from f- \lt(f)$
    \EndIf
    \Until{$f = 0$}
    \Return\{\texttt{quotient}, \texttt{reminder}\}
    \EndProcedure
  \end{algorithmic}
\end{algorithm}

At the end of this algorithm, we can write
\[
f =\left(\sum_{\mathrm{finite}}  h_i\cdot g_i\right) + r.
\]
If the reminder is zero, then $f\in I$. We can implement this
algorithm (code from Greg Smith)
\begin{macaulay2}
division = (f,G) -> (
S := ring f;
p := f;
r := 0_S;
m := #G;
Q := new MutableHashTable;
for j from 0 to m-1 do Q#j = 0_S;
while p != 0 do (
i := 0;
while i < m and leadTerm(p) % leadTerm(G#i) != 0 do i = i+1;
if i < m then (Q#i = Q#i + (leadTerm(p) // leadTerm(G#i));
p = p - (leadTerm(p) // leadTerm(G#i)*G#i))else (r = r + leadTerm(p);
p = p - leadTerm(p)));
L := apply(m, j -> Q#j);
return (r,L));
\end{macaulay2}
%% I think f % matrix{{g_1},{g_2}}
%% should give the reminder, but it doesn't work.


However, this algorithm is
insufficient. Consider an example from \cite{hS2003} where
\begin{align*}
  f &= x^2 - y^2\\
  g_1 &= x^2 + y\\
  g_2 &=xy+x \\
  I &= (g_1,g_2).
\end{align*}
Let's run the division algorithm manually in \macaulay and try to see
if we what happens.

\begin{macaulay2}
R = ZZ/101[x,y];
f = x^2-y^2;
g1 = x^2+y;
g2 = x*y+x;
leadTerm f
leadTerm g1
g%g1
\end{macaulay2}

Using the monomial order \texttt{GRevLex}, $\lt(f) = x^2$ and
$\lt(g_1) = x^2$. Hence, we divide $f$ by $g_1$ and compute the
remainder, which comes out to $-y^2-y$. Since neither $\lt(g_1) =
x^2$ nor $\lt(g_2) = xy$ divide $y^2$, the algorithm concludes with
\[
f  = 1\cdot g_1 + \underbrace{(-y^2-y)}_{\text{remainder}}.
\]
this would seem to indicate that $F\notin I$. However,
\[
f = x g_2 - y g_1.
\]
so in fact, $f\in I$. Our algorithm didn't work because the lead terms
of $g_1$ and $g_2$ cancelled. To remedy this, we will enlarge our set
of generators to include more lead terms. This leads us to the
definition of a Gr\"obner basis.

\begin{definition}
  A subset $\{g_1,\dots, g_m\}\subset I$ of an ideal is a
  \dfn{Gr\"obner basis} for $I$ if
  \[
  \lt(I) = (\lt(g_1),\dots, \lt(g_m)).
  \]
\end{definition}

\begin{proposition}
  Suppose $(g_1,\dots,g_m)$ is a Gr\"obner basis for $I$. Then
  \[
  I = (g_1,\dots,g_m).
  \]
  \begin{proof}
    Suppose $f\in I$. Using the division algorithm on the ideal
    $\{g_1,\dots,g_m\}$, there will always be a $g_i$ such that
    $\lt(g_i)|\lt(f)$. Hence $f$ can be expressed as a finite sum
    \[
    f = \sum_\mathrm{finite} h_i g_i.
    \]
  \end{proof}
\end{proposition}

Now suppose that $\{g_1,\dots,g_m\}$ is a Gr\"obner basis for some
ideal. We can then write:
\[
0 \to S \to k[\x]^n \to k[\x]/I \to 0
\]
$S$ is a syzygy, the kernel of this map of free modules.

\begin{definition}
  Let $f,g\in k[\x]$ be nonzero polynomials. The
  \dfn{syzygy-polynomial} is given by
  \[
  S(f,g) := \frac{\lcm(\lm(f),\lm(g))}{\lt(f)}\cdot f - \frac{\lcm(\lm(f),\lm(g))}{\lt(g)}\cdot g.
  \]
\end{definition}
We can define the syzygy-polynomial in \macaulay.
\begin{macaulay2}
S = (g1,g2) ->
lcm(leadMonomial(g1),leadMonomial(g2))//leadTerm(g1)*g1-
lcm(leadMonomial(g1),leadMonomial(g2))//leadTerm(g2)*g2;
R = ZZ/101[x,y];
g1 = x^2+y;
g2 = x*y+x;
S(g1,g2)
\end{macaulay2}






\begin{theorem}[Buchberger's criterion]
  A set $G=\{g_1,\dots,g_n\}\subset k[\x]$ is Gr\"obner basis if and
  only if for all $i\ne j$ 
  \[
  S(g_i,g_j) \equiv 0 \pmod{G}.
  \]
  In this case $g_i$ and $g_j$ are called an \dfn{s-pair} or a
  $\dfn{syzygy-pair}$.
  \begin{proof}
    $(\Rightarrow)$ Suppose that $\{g_1,\dots,g_m\}$ is a Gr\"obner
    basis. Since $S(g_i,g_j)\in (g_1,\dots,g_m)$ we are done.

    $(\Leftarrow)$ Suppose that $f\in (g_1,\dots,g_m)$ and that for
    all $i\ne j$
    \[
    S(g_i,g_j) \equiv 0 \pmod{G}.
    \]
    We must show that
    \[
    \lt(f) \in (\lt(g_1),\dots, \lt(g_n)).
    \]
    Since $f\in (g_1,\dots,g_m)$, we know
    \[
    f = \sum_{\mathrm{finite}} h_i g_i.
    \]
    At this point there are two cases, either $\lm(f) = \lm(g_i)$, in
    which case we are done. The second case when $\lm(f) \ne \lm(g_i)$
    requires attention. We refer the interested reader
    to~\cite{CLO2007}.
  \end{proof}
\end{theorem}


So here's the basic idea: Take your favorite ideal, compute all the
s-pairs, add them to your idea. If you do this enough, you'll satisfy
Buchberger's criterion.


\begin{algorithm}[Buchberger algorithm]
  \hfill
  \begin{algorithmic}[1]
    \Procedure{Gr\"obner basis}{$f_1,\dots,f_m$} 
    \State $G = \{f_1,\dots,f_m\}$
    \Repeat
    \State $G':=G$
    \For{each pair $p,q\in G'$} $S:= S(p,q)$
    \If{$S\not\equiv 0 \pmod{g\in G}$}
    $G\from G\cup\{S\}$
    \EndIf
    \EndFor
    \Until{$G' = G$}
    \EndProcedure
  \end{algorithmic}
\end{algorithm}

Let's see it in action with \macaulay.

\begin{macaulay2}
division = (f,G) -> (
S := ring f;
p := f;
r := 0_S;
m := #G;
Q := new MutableHashTable;
for j from 0 to m-1 do Q#j = 0_S;
while p != 0 do (
i := 0;
while i < m and leadTerm(p) % leadTerm(G#i) != 0 do i = i+1;
if i < m then (Q#i = Q#i + (leadTerm(p) // leadTerm(G#i));
p = p - (leadTerm(p) // leadTerm(G#i)*G#i))else (r = r + leadTerm(p);
p = p - leadTerm(p)));
L := apply(m, j -> Q#j);
return (r,L));

S = (g1,g2) ->
lcm(leadMonomial(g1),leadMonomial(g2))/leadTerm(g1)*g1-
lcm(leadMonomial(g1),leadMonomial(g2))/leadTerm(g2)*g2;

R = ZZ/101[x,y];
g_1 = x^2+y;
g_2 = x*y+x;
s = S(g_1,g_2)
division(s,{g_1,g_2})
g_3=s
s = S(g_1,g_3)
division(s,{g_1,g_2,g_3})
g_4=s
s = S(g_1,g_4)
division(s,{g_1,g_2,g_3,g_4})
s = S(g_2,g_3)
division(s,{g_1,g_2,g_3,g_4})
s = S(g_2,g_4)
division(s,{g_1,g_2,g_3,g_4})
s = S(g_3,g_4)
division(s,{g_1,g_2,g_3,g_4})
ideal (g_1,g_2,g_3,g_4)
gens gb ideal (g_1,g_2)
\end{macaulay2}

At this point, we can solve the ideal memebership problem.

\begin{proposition}
  Let $G = \{g_1,\dots,g_m\}$ be a Gr\"obner basis for and ideal
  $I\subset k[x_1,\dots,x_n]$. An element $f\in I$ if and only if the
  reminder of $f$ by $G$ is zero.
  \begin{proof}
    $(\Rightarrow)$ If $f\in I$, then, since $I$ is the ideal
    generated by $G$, the reminder is zero.

    $(\Leftarrow)$ If the reminder of $f$ is zero when divided by $G$,
    then, $f\in I$.
  \end{proof}
\end{proposition}

Moreover, if we are working in a ring $R=k[x_1,\dots,x_n]$ and we wish
to know if an element is in the radical of an ideal, we can use the
``Rabinowitch trick'' \index{Rabinowitch trick} and we claim $a\in
\rad{I}$ if and only if
\[
1 \in I + (1-ya) \subset R[y]
\]
Note, we know that given a ring, $R/I$, in $(R/I)[y]$, the Jacobson
radical equals the nilradical. Hence if $y$ is in the radical of $I$
if and only if $(1-ya)$ is a unit in $(R/I)[y]$.

\end{document}



