\documentclass{ximera}

\input{../preamble.tex}

\title{Monomials and orderings}

\begin{document}
\begin{abstract}
  We set up definitions and notation for the study of Groebner
  bases. Sources and references: \cite{CLO2007}.
\end{abstract}
\maketitle

Throughout this chapter our ring will be $R=k[x_1,\dots,x_n] =
k[\x]$. We will denote the whole numbers by $\omega =
\{0,1,2,3,\dots\}$.


\begin{definition}
  A \dfn{monomial} in $k[x_1,\dots,x_n]$ is a product of the form
  \[
  \x^{\valpha} = x_1^{\alpha_1}\cdot x_2^{\alpha_2}  \cdots x_n^{\alpha_n}.
  \]
  where $\valpha = (\alpha_1,\dots,\alpha_n) \in
  \omega^n$.  A \dfn{polynomial} is finite sum of monomials with
  coefficients $c_\alpha\in k$,
  \[
  F = \sum_{\mathrm{finite}} c_{\valpha} \x^{\valpha}.
  \]
\end{definition}


\begin{definition}
  Let $R = k[\x]$ and consider a polynomial
  \[
  F =  \sum_{\mathrm{finite}} c_{\valpha} \x^{\valpha}.
  \]
  \begin{itemize}
    \item The \dfn{degree} of a monomial $\x^{\valpha}$ is
      $\valpha=(\alpha_1,\dots,\alpha_n)\in \omega^n$.
    \item The \dfn{total degree} of a monomial $\x^{\valpha}$ is
      $|\boldsymbol{\alpha}| = \sum_{i=1}^n \alpha_i$.
    \item Each $c_{\valpha}$ is the \dfn{coefficient} of the
      monomial $\x^{\valpha}$.
    \item If $c_{\valpha} \ne 0$, then
      $c_{\valpha}\x^{\valpha}$ is a \dfn{term} of
      $F$.
    \item The \dfn{total degree} of $F$, is the maximum total degree
      of the monomials in $F$.
  \end{itemize}
\end{definition}



\begin{definition}
  A \dfn{monomial ordering} $(k[\x],>)$ is a relation on the set of
  monomials in $k[\x]$ such that:
  \begin{enumerate}
  \item If $\x^{\valpha}$ and $\x^{\vbeta}$ are
    monomials in $R$, then exactly one of the three statements
      \[
      \x^{\valpha} > \x^{\vbeta},\quad \x^{\valpha} = \x^{\vbeta},
      \quad \x^{\vbeta} > \x^{\valpha}.
      \]
      holds. This means that ordering $>$ is a total ordering.
    \item If $\x^{\valpha}>\x^{\vbeta}$ and
      $\vgamma\in\omega^n$, then
      $\x^{\valpha+\vgamma}>\x^{\vbeta+\vgamma}$.
    \item Every nonempty subset of the set of monomials of $k[\x]$ has
      a least element with respect to $>$. This means that the
      monomials of $k[\x]$ are well-ordered.
  \end{enumerate}
\end{definition}

Now let's look at some explicit monomial orderings.


\begin{definition}[Lexicographic order]\index{lexicographic order}
  Consider monomials $\x^\valpha,\x^\vbeta\in k[\x]$. We say
  \[
  \x^\valpha \underset{\mathtt{Lex}}{>}\x^\vbeta
  \]
  when the left-most nonzero entry of $\valpha-\vbeta$ is positive.
\end{definition}

The computer algebra system \macaulay\ will write the terms of a
polynomial in decreasing order relative to the specified monomial
order.

\begin{macaulay2}
ZZ/101[x,y,MonomialOrder=>Lex];
y^3+ x*y + x + y + 1
\end{macaulay2}

Here we see:
\[
xy \underset{\mathtt{Lex}}{>} x \underset{\mathtt{Lex}}{>} y^3
\underset{\mathtt{Lex}}{>} y \underset{\mathtt{Lex}}{>} 1.
\]



\begin{definition}[Graded reverse lexicographic order]\index{graded reverse lexicographic order}
  Consider monomials $\x^\valpha,\x^\vbeta\in k[\x]$. We say
  \[
  \x^\valpha \underset{\mathtt{GRevLex}}{>}\x^\vbeta
  \]
  when $|\valpha| > |\vbeta|$ or $|\valpha| = |\vbeta|$ and the
  right-most nonzero entry of $\valpha-\vbeta$ is negative.
\end{definition}

The order $\mathtt{GRevLex}$ is the default ordering in \macaulay.

\begin{macaulay2}
ZZ/101[x,y]
y^3+ x*y + x + y + 1
\end{macaulay2}

Here we see:
\[
y^3 \underset{\mathtt{GRevLex}}{>} xy \underset{\mathtt{GRevLex}}{>} x  \underset{\mathtt{GRevLex}}{>} y \underset{\mathtt{GRevLex}}{>} 1
\]

\begin{definition}[Weight order]\index{weight order}
  Consider monomials $\x^\valpha,\x^\vbeta\in k[x_1,\dots, x_n]$ and
  let $\vec{w}$ be a vector in $\Z^n$. We say
  \[
  \x^\valpha \underset{\mathtt{w}}{>}\x^\vbeta
  \]
  when $\vec{w}\dotp \valpha > \vec{w}\dotp \vbeta$ or $\vec{w}\dotp
  \valpha = \vec{w}\dotp \vbeta$ and $\x^\valpha
  \underset{\mathtt{GRevLex}}{>}\x^\vbeta$.
\end{definition}

Weighted orderings are quite versatile. Here is a basic example with
$\vec{w} = (0,3)$ in \macaulay.

\begin{macaulay2}
ZZ/101[x,y,MonomialOrder=>{Weights=>{0,3}}]
y^3+ x*y + x + y + 1
\end{macaulay2}

Here we see:
\[
y^3 \underset{\mathtt{w}}{>} xy \underset{\mathtt{w}}{>} y \underset{\mathtt{w}}{>} x \underset{\mathtt{w}}{>} 1
\]

\begin{proposition}
  Each of the orderings above defines a monomial ordering.
  \begin{sketch}
    Just check the conditions.
  \end{sketch}
\end{proposition}



\end{document}
