\documentclass{ximera}

\input{../preamble}

\author{Bart Snapp}

\title{Going-down}

\begin{document}
\begin{abstract}
  We analyze the going-down property. Sources and references:
  \cite{AM1969}.
\end{abstract}
\maketitle

The going-down property also is closely related to integral
extensions. Let's recall the going-down property:
\begin{description}
\item[Going-down]\index{going-down} If $\p_1,\p_2\in\spec(R)$ with $\p_1\subset\p_2$ and
  $P_2\cap R = \p_2$, then there exists $P_1$ with $P_1\subset P_2$
  and $P_1\cap R = \p_1$.
  \[
  \begin{tikzcd}[row sep=0em,column sep=0em]
    R    & \subset & S \\
    \cup &         & \cup \\
    \p_2 & \subset & P_2 \\
    \cup &         & \cup \\
    \p_1 & \subset & \exists P_1  
  \end{tikzcd}
  \]
\end{description}


We will strengthen Proposition~\ref{P:localint}.


\begin{proposition}[Integral closure and localization]\label{P:localintS}
  Let $R\subset S$ with $A$ being the integral closure of $R$
  in $S$. If $U\subset R$ is a multiplicatively closed set, then
  $U^{-1}A$ is the integral closure of $U^{-1}R$ in $U^{-1}S$.
  \begin{proof}
    First note that by Proposition~\ref{P:localint}, $U^{-1} A$ is
    integral over $U^{-1} R$. Now suppose there is an element $s/u\in
    U^{-1} S$, that is integral over $R$.  We must show that this
    element is in $S^{-1} A$. Consider the integral dependence
    equation for $s/u$ over $S^{-1} R$
    \[
    (s/u)^{n} + \left(r_{n-1}/u_{n-1}\right)(s/u)^{n-1}+\cdots +  \left(r_{1}/u_{1}\right)(s/u) + r_{0}/u_{0} = 0
    \]
    where $r_i\in R$ and $u_i\in U$. Clear denominators, by
    multiplying by $(uu_{n-1}\cdots u_0)^n$, then this is the integral
    dependence equation for $su_{n-1}\cdots u_0$. Since $A$ is the
    integral closure of $R$, we see that
    \[
    su_{n-1}\cdots u_0\in A
    \]
    and hence $s/u\in U^{-1} A$.
  \end{proof}
\end{proposition}


\begin{definition}
  Let $A$ be a domain. The \dfn{integral closure} of $A$, without
  reference to another ring, means the integral closure of $A$ in its
  field of fractions $\ff(A)$. In this case we write $\tilde{A}$ for
  the integral closure.
\end{definition}



\begin{proposition}
  Let $A$ be an integral domain, the following are equivalent:
  \begin{enumerate}
  \item $A$ is integrally closed.
  \item $A_\p$ is integrally closed for all $\p\in\spec(A)$.
  \item $A_\m$ is integrally closed for all maximal ideals.
  \end{enumerate}
  \begin{proof}
    (a)$\Rightarrow$(b)$\wedge$(c) Follows from Proposition~\ref{P:localintS}.

    (b)$\vee$(c)$\Rightarrow$(a) Consider $\iota: A\into\tilde{A}$ via
    the identity map. $A$ is integrally closed if and only if $\iota$
    is surjective. However, if $A_\p\into\tilde{A}_\p$ is surjective
    for all prime ideals (similarly maximal ideals), then
    $A\into\tilde{A}$ is surjective.
  \end{proof}
\end{proposition}


\begin{definition}
   Let $R\subset S$ with $I\subset R$ an ideal. The \dfn{integral
     closure of an ideal} $I$ in $S$ is the set of all elements of $S$
   that are integral over $I$.
\end{definition}


\begin{proposition}\label{P:intideal}
  Let $R\subset A\subset S$ be rings with $A$ being the integral
  closure of $R$ in $S$. Let $I$ be an ideal of $R$. Then
  \[
  \{\text{integral closure of $I$ in $S$}\} = \rad{IA}.
  \]
  \begin{proof}
    $(\subset)$ Suppose that $x$ is integral over $I$. Then
    \[
    x^n + a_{n-1}x^{n-1} + \dots + a_1 x + a_0 = 0
    \]
    for $a_i\in I$. Hence, $x\in A$, and $x^n\in IA$, so $x\in
    \rad{IA}$.

    $(\supset)$ Suppose that $x\in\rad{IA}$. This means that
    \[
    x^n = \sum_{i=1}^m a_i x_i
    \]
    where $a_i\in I$ and $x_i\in A$. Set $M = R[x_1,\dots,
      x_m]$. Since each $x_i$ is integral over $R$, $M$ is finitely
    generated over $R$. Now, $x^n M \subset I M$, hence by the
    determinant trick (Proposition~\ref{P:determinanttrick}) where
    $\phi$ is multiplication by $x^n$, we have that $x^n$ is integral
    over $R$, and we are done.
  \end{proof}
\end{proposition}


\begin{proposition}\label{P:intff}
  Let $A\subset B$ be integral domains with $A = \tilde{A}$. Let $b\in
  B$ be integral over an ideal $I\subset A$.
  \begin{enumerate}
  \item The element $b$ is algebraic over $\ff(A)$.
  \item If $x^n + a_{n-1}x^{n-1} + \dots + a_1 x + a_0$ is a minimal
    polynomial for $b$ over $\ff(A)$, then $a_i\in \rad{I}$.
  \end{enumerate}
  \begin{proof}
    It follows directly that $b$ is algebraic over $\ff(A)$.

    Let $\ff(A)\subset K$ an extension field containing all of the
    conjugates of $x$, call them $x_1,\dots, x_n$. Each $x_i$ is
    integral over $A$ via the same integral dependence equation. The
    coefficients of the integral dependence equation are polynomials
    in the $x_i$. By Proposition~\ref{P:intideal}, they are integral
    over $I$, and are in $\rad{I}$.
  \end{proof}
\end{proposition}


\begin{theorem}[Going-down]
  If $A\subset B$ are integral domains, $A=\tilde{A}$, and $B$ is
  integral over $A$, then the going-down property holds.
  \begin{proof}
    We have the following setup:
    \[
    \begin{tikzcd}[row sep=0em,column sep=0em]
      A=\tilde{A}    & \subset & B \\
      \cup &         & \cup \\
      \p_2 & \subset & P_2 \\
      \cup &         & \cup \\
      \p_1 & \subset & \exists P_1  
    \end{tikzcd}
    \]
    We claim that $P_1 = \p_1 B_{P_2}$. Note, by construction we have that
    \[
    \p_1 \subset \p_1 B_{P_2}\cap A,
    \]
    hence we must show that if $x\in \p_1 B_{P_2}\cap A$, then
    $x\in\p_1$.

    Consider $x\in \p_1 B_{P_2}$. This means that $x = y/u$ where
    $u\in B-P_2$ and $y\in \p_1 B$. By Proposition~\ref{P:intideal},
    $y$ is integral over $\p_1$, and by Proposition~\ref{P:intff}, its
    minimal polynomial over $\ff(A)$ is
    \[
    y^n + p_{n-1}y^{n-1} + \dots + p_1 y + p_0 = 0
    \]
    where each $p_i\in\p_1$.

    Suppose further that $x\in \p_1 B_{P_2}\cap A$. Now since $x =
    y/u$, we have $u = y/x$, where $x\in\ff(A)$. Now the minimal
    polynomial for $u$ can be found by dividing the polynomial above
    by $x^n$ to get
    \[
    u^n + q_{n-1} u^{n-1} + \dots + q_1 u + q_0 = 0
    \]
    where $q_i = p_i/x^i$. Thus
    \[
    x^i q_i = p_i \in \p_1.
    \]
    But $u$ is integral over $A$, so each $q_i\in A$. Now suppose that
    $x\notin\p_1$, then the equation immediately above shows that
    $q_i\in\p_1$, and so the equation above that shows
    \[
    u^n\in \p_1 B \subset \p_2 B \subset P_2
    \]
    and so $u\in P_2$, a contradiction. Hence $x\in \p_1$, and so
    $\p_1 B_{P_2}\cap A = \p_1$.
  \end{proof}
\end{theorem}



\begin{corollary}
  Let $R\subset S$ have the going-down property. If $\p\in\spec{R}$
  and $P\in\spec(S)$, then $\h(P)= \h(\p)$.
\end{corollary}



For some interesting extra reading check out:
\begin{itemize}
\item \link[\textit{Prime ideals and integral dependence}, I.S.\ Cohen and A.\ Seidenberg, Bulletin of the American Mathematical Society, 52, 1946, 252--261]{https://doi-org.proxy.lib.ohio-state.edu/10.1090/S0002-9904-1946-08552-3}.
\end{itemize}


\end{document}
