\documentclass{ximera}

\input{../preamble.tex}

\author{Bart Snapp}

\title{Applications of primary decomposition}

\begin{document}
\begin{abstract}
  We give examples and applications of primary decomposition. Sources
  and references: \cite{AM1969,sD2008,dE1995}.
\end{abstract}
\maketitle



\begin{definition}
  Let $R$ be as ring. An ideal $\q$ is called \dfn{primary}, if $\q\ne
  R$ and
  \[
  xy\in\q \quad \Rightarrow \quad x\in\q\text{ or }y^m\in\q.
  \]
\end{definition}

\begin{exercise}
  Let $R$ be a Noetherian ring and $\q$ be an ideal of $R$. Prove that
  $\q$ is $\p$-primary if and only if $\ass(R/I) = \{\p\}$.
\end{exercise}

\begin{exercise}
  Show that every prime ideal is primary.
\end{exercise}

\begin{exercise}
  Show that if an idea $\q$ is $\p$-primary, then $\rad{\q} = \p$.
\end{exercise}

\begin{exercise}
  Let $k$ be a field, and consider $R=k[x,y,z]/(z^2-xy)$. Let $\p = (\bar{x},\bar{y})$.
  \begin{enumerate}
  \item Show that $\p$ is a prime ideal in $R$.
  \item Show that $\p^2$ is not $\p$-primary.
  \end{enumerate}
\end{exercise}

\begin{exercise}
  Let $R$ be a UFD and $\p = (p)$ where $p$ is a prime element. Show
  that $\p^n$ is a $\p$-primary ideal for all $n>0$.
\end{exercise}

\begin{exercise}
  Prove that if $\a$ is $\p$-primary and $\b$ is $\q$-primary, then
  $\a$ and $\b$ are comaximal.
\end{exercise}



\begin{theorem}[Primary decomposition of ideals]
  Let $R$ be a Noetherian ring, and $I$ be an ideal of $R$. We may write
  \[
  I = \bigcap_{i=1}^n \q_i
  \]
  uniquely and minimally where each $\q$ is \dfn{primary}
  \begin{proof}
    Let $M = R/I$. By Primary decomposition (Theorem~\ref{T:primarydecomp}),
    \[
    I = \bigcap_{i=1}^n \q_i
    \]
    where for each $i$, $\ass(R/\q_i) = \{\p_i\}$, if $i\ne j$, then
    $\p_i\ne\p_j$, and $\ass(R/I) = \{\p_1,\dots,\p_n\}$.
  \end{proof}
\end{theorem}

\begin{theorem}[Fundamental theorem of arithmetic]
  Every integer greater than $1$ can be factored uniquely (up to
  ordering) into primes.
  \begin{proof}
    Let $a\in\Z$. Consider the ideal generated by $a$. From above
    \[
    a\Z = \bigcap_{i=1}^n \q_i
    \]
    where $\ass(\Z/\q_i) = p_i\Z$ and $\ass(\Z/a\Z) = \{p_1\Z,\dots,
    p_n\Z\}$. Since the prime ideas of $\Z$ are principally generated,
    and the primary ideals of $\Z$ are simply prime powers, we are
    done.
  \end{proof}
\end{theorem}

\begin{exercise}%hM1986 page 161
   A Noetherian integral domain $R$ is a UFD if and only if every minimal
   prime ideal is principal.
\end{exercise}

\begin{theorem}[Fundamental theorem of finitely generated abelian groups]
  Every finitely generated Abelian group $G$ is isomorphic to a
  direct sum of cyclic groups of prime-power order and infinite cyclic
  groups,
  \[
  G \iso \Z^n \dsum \directsum_{i=1}^{m} \Z_{p_i^{k_i}},
  \]
  where the $p_i$ are not necessarily distinct prime numbers.
  \begin{proof}
    A finitely generated Abelian group is a $\Z$-module. Consider
    \[
    0 \to \syz_1(G) \to \Z^n \to G \to 0 
    \]
    where $\syz_1(G)$ is the first syzygy\index{syzygy} module, meaning $\syz_1(G) =
    \ker(\Z^n \to G)$. By Primary decomposition (Theorem~\ref{T:primarydecomp}),
    \[
    \syz_1(G) = \bigcap_{i=1}^n N_i
    \]
    where $\ass(\Z^n/N_i) = \{\p_i\}$ and $\ass(G) =
    \{p_1\Z,\dots,p_n\Z\}$. If $\ass(\Z^n/N_i) = \{p_i\Z\}$, this
    means that $N_i = 0$ or $N_i = (\Z/p_i\Z)^j$.
  \end{proof}
\end{theorem}


\begin{theorem}[Primary decomposition and localization]
  Consider a Noetherian ring $R$, a multiplicatively closed subset of
  $S\subset R$, a finitely generated $R$-module $M$, and an
  $R$-submodule $N\subset M$, with a primary decomposition. We may
  renumber the factors $N_i$ so that for $i=1,\dots, m$, $\p_i\cap S
  =\emptyset$, and if $i=m+1,\dots, n$, $\p_i\cap S \ne\emptyset$,
  \[
  S^{-1}N = \bigcap_{i=m}^n S^{-1}N_i
  \]
  where for each $i$, $\ass(S^{-1}(M/N_i))$ contains exactly one prime ideal
  of $S^{-1}R$ and if $i\ne j$, then $\ass(S^{-1}(M/N_i)) \ne \ass(S^{-1}(M/N_j))$.
  \begin{sketch}
    Apply Proposition~\ref{P:assassinlocal}, to Theorem~\ref{T:primarydecomp}.
  \end{sketch}
\end{theorem}




\begin{theorem}[Decomposing varieties]
  Let $V\subset\A^n$ be an affine algebraic variety over an
  algebraically closed field. We may express
  \[
  V = \bigcup_{i=1}^n U_i
  \]
  uniquely and minimally where each $U_i$ is an irreducible variety.
\end{theorem}


\begin{exercise}
  Let $R$ be a Noetherian ring. Prove that $\spec(R)$ is a Noetherian topological space. 
\end{exercise}


Let's see some examples in \macaulay.

\begin{macaulay2}
R = QQ[x,y];
I = ideal (x^2,x*y);
primaryDecomposition I
ass I
\end{macaulay2}

Note, when write \texttt{ass I} in \macaulay\, it is computing
$\ass(R/I)$.

\begin{macaulay2}
R = QQ[x,y,z];
I = ideal(x^2,x*y,x*z);
primaryDecomposition I
ass I
\end{macaulay2}




%% \subsection{Induced map}


%% answer question:

%%   Given a morphism $\Phi\supp(M) \to \supp(N)$, can you induce
%%   \[
%%   \Phi^\sharp : N\to M?
%%   \]


For some interesting extra reading check out:
\begin{itemize}
\item \link[\textit{Direct methods for primary decomposition}, D.\ Eisenbud, C.\ Huneke, W.\ Vasconcelos, Inventiones mathematicae, December 1992]{https://link.springer.com/content/pdf/10.1007\%2FBF01231331.pdf}.

\item \link[\textit{The question of finitely many steps in polynomial ideal theory}, G.\ Hermann, ACM SIGSAM Bulletin, September 1995]{https://dl.acm.org/ft_gateway.cfm?id=307342&ftid=82112&dwn=1&CFID=78879703&CFTOKEN=a9e554efb35114eb-60020924-030F-8CA8-95F3344136DF695C}.


\end{itemize}



\end{document}
