\documentclass{ximera}

\input{../preamble.tex}

\author{Bart Snapp}

\title{Associated primes are assassins}

\begin{document}
\begin{abstract}
  We introduce associated primes, also known as assassins. Sources and
  references: \cite{sD2008, dE1995, mR1995,jpS2000}.
\end{abstract}
\maketitle

\begin{definition}
  Let $R$ be a ring and $Q$ be a $R$-module. The \dfn{assassins} of
  $Q$, denoted by $\ass(Q)$, is a set such that:
  \begin{itemize}
  \item $\ass(Q)\subset\spec(R)$.
  \item For each $\p\in\ass(Q)$, $Q$ contains a submodule isomorphic
    to $R/\p$.
  \end{itemize}
  The set of prime ideals in $\ass(Q)$ is often called the
  \dfn{associated primes} of $Q$.
\end{definition}

\begin{exercise}
  Let $Q$ be an $R$-module. Show there exists $x\in Q$ such that
  $\ann(x) = \p$ if and only if $Q$ contains a submodule isomorphic to
  $R/\p$. This explains why one might call $\p$ an assassin of $Q$.
\end{exercise}

\begin{exercise}
  Consider $\Z_6$ as a $\Z$-module. Find the assassins of $\Z_6$.
\end{exercise}

\begin{exercise}
  Let $k$ be a field, consider the $k[x]$-module $k[x]/(x^5)$. Find
  the assassins of $k[x]/(x^5)$.
\end{exercise}

\begin{exercise}
  Let $k$ be a field, consider the $k[x,y]$-module $k[x,y]/(xy)$. Find
  the assassins of $k[x,y]/(xy)$.
\end{exercise}




\begin{proposition}
  Let $R$ be a Noetherian ring, and $Q$ be a finitely generated
  $R$-module. Consider
  \[
  \mathcal{S} = \{\ann(x): x\in Q\text{ and } x\ne0\}.
  \]
  The maximal elements of $\mathcal{S}$ are prime.
  \begin{proof}
    Since $R$ is Noetherian, and $\mathcal{S}$ is nonempty, it
    contains $(0)$, $\mathcal{S}$ contains a maximal element, call it
    $\p$.  Let $x$ be a nonzero element of $Q$ such that $\ann(x) =
    \p$.  Suppose $ab\in\p$, and $a\notin\p$.  In this case,
    $\p=\ann(x) \subset \ann(ax)$. Since $\p$ is maximal, $\p =
    \ann(ax)$. However, $b\in\ann(ax) = \p$, hence $\p$ is prime.
  \end{proof}
\end{proposition}

To quote David Eisenbud, ``ideals maximal with respect to some
property have an uncanny tendency to be prime.''


\begin{corollary}[Assassins are nonempty]
  Let $R$ be a Noetherian ring, and $Q$ be a finitely generated
  $R$-module. If $Q\ne 0$, then $\ass(Q) \ne \emptyset$.
  \begin{proof}
    Since $\mathcal{S}$ from above is nonempty, it contains maximal
    elements and these elements are prime.
  \end{proof}
\end{corollary}




\begin{example}
  Note if $R$ is not Noetherian, it may be the case that
  $\ass(M) = \emptyset$. Consider $R = k[x_1,x_2,\dots]$. Let $I$ be the ideal generated by
  \[
  \{x_i^{i+1}:i\ge 1\}.
  \]
  Finally, let $M = R/I$. In this case, $\supp(M) =
  \{(x_1,x_2,\dots)\}$; however, $\ass(M) = \emptyset$.
\end{example}




\begin{corollary}
  If $R$ is a Noetherian ring and $Q$ is a finitely generated
  $R$-module, then there is a finite filtration
  \[
  0 = Q_0 \subsetneq Q_1 \subsetneq \dots \subsetneq Q_n = Q 
  \]
  where $Q_i/Q_{i-1} \iso R/\p_i$, where $\p_i\in\spec(R)$.
  \begin{proof}
    By the corollary above, $Q$ contains a submodule isomorphic to
    $R/\p_1$. This submodule will be $Q_1$. If $Q = Q_1$, we are
    done. Proceed inductively, suppose that we can prove our statement
    for up to $n$ submodules. If $Q \ne Q_n$, then $Q/Q_n$ contains a
    submodule isomorphic to $R/\p_{n+1}$, call it $Q_{n+1}/Q_n$. Since
    $Q$ is Noetherian, this ascending chain stabilizes.
  \end{proof}
\end{corollary}

\begin{exercise}
  Let $R$ be a Noetherian ring and $M$ be a finitely generated $R$-module. Consider the canonical map
  \[
  M\to \prod_{\p\in\ass(M)} M_\p.
  \]
  Prove that this map is injective.
\end{exercise}



\begin{lemma}[Irreducibility and assassins]\label{L:irrassassin}
  Let $R$ be a Noetherian ring and $M$ be finitely generated
  $R$-module. If $N$ is an irreducible submodule of $M$, then
  $\ass(M/N)=\{\p\}$.
  \begin{proof}
    Let $N$ be an irreducible submodule of $M$. Seeking a
    contradiction, suppose that $\ass(M/N) = \{\p,\q\}$. Then there
    are elements $\bar{x},\bar{y}\in M/N$ such that $\p=\ann(\bar{x})$
    and $\q=\ann(\bar{y})$. We claim
    \[
    \bar{x}R \cap \bar{y}R = 0.
    \]
    To see this, consider $a\in\bar{x}R \cap \bar{y}R$, $r,s\in R$,
    $p\in \p-\q$, and write
    \begin{alignat*}{2}
      a &= \bar{x} r & &= \bar{y} s\\
      pa &= p\bar{x} r & &= p\bar{y} s\\
      0  &= 0  & &= p\bar{y} s.
    \end{alignat*}
    This means that $ps\in\q$, which implies $s\in\q$. Hence
    \[
    a = \bar{y}s = 0.
    \]
    Since $\bar{x}R \cap \bar{y}R = 0$, this means that $(xR
    +N)\cap(yR+N) = N$, contradicting the irreducibility of $N$.
  \end{proof}
\end{lemma}

\begin{proposition}[Localization and assassins]\label{P:assassinlocal}
  Let $R$ be a Noetherian ring, $S$ be a multiplicatively closed
  subset of $R$, $\p\in\spec(R)$ with $\p\cap S = \emptyset$, and $Q$
  be an $R$-module. Then
  \[
  \p \in \ass_R(Q) \Leftrightarrow S^{-1}\p\in \ass_{S^{-1}R}(S^{-1}Q)
  \]
  whenever $S\cap\p = \emptyset$.
  \begin{proof}
    $(\Rightarrow)$ If $\p\in\ass_R(Q)$, then for some $x\in Q$
    \begin{align*}
    x\cdot p &= 0\\
    \frac{x}{1} \cdot \frac{p}{1} &= 0
    \end{align*}
    so $S^{-1} \p\in \ass(S^{-1}Q)$.

    $(\Leftarrow)$ If $S^{-1}\p\in \ass_{S^{-1}R}(S^{-1}Q)$, then
    there is an injective homomorphism
    \[
    \iota:\frac{S^{-1}R}{S^{-1}\p}\into S^{-1}Q
    \]
    Since $\p$ is finitely generated (see \cite[Proposition 2.10]{dE1995})
    \[
    \hom_{S^{-1}R}\left(\frac{S^{-1}R}{S^{-1}\p},S^{-1}Q\right) = S^{-1}\hom_{R}(R/\p,Q)
    \]
    Since $S\cap\p=\emptyset$ we can clear denominators of $\iota$ to find the
    injection we are looking for.
  \end{proof}
\end{proposition}





\begin{theorem}[Assassins are minimal]\label{T:AssassinsMinimal}
  If $R$ is a Noetherian ring and $Q$ is a finitely generated
  $R$-module, consider the finite filtration
  \[
  0 = Q_0 \subsetneq Q_1 \subsetneq \dots \subsetneq Q_n = Q 
  \]
  where $Q_i/Q_{i-1} \iso R/\p_i$. In this case
  \[
  \ass(Q) \subset \{\p_1,\dots,\p_n\}\subset \supp(Q)
  \]
  and each of these sets have the same minimal elements.
  \begin{proof}
    First we'll show that $\{\p_1,\dots,\p_n\}\subset \supp(Q)$ and
    that these sets have the same minimal elements. Localize the
    finite filtration
    \[
    0 = (Q_0)_\p \subsetneq (Q_1)_\p \subsetneq \dots \subsetneq (Q_n)_\p = Q_\p,
    \]
    with the factors being
    \begin{align*}
      (Q_i)_\p/(Q_{i-1})_\p &\iso \left(Q_i/Q_{i-1}\right)_\p\\
      &\iso \left(R/\p_i\right)_\p.
    \end{align*}
    Hence, $Q_\p \ne 0$ if and only if there is $\p_i$ such that
    $\left(R/\p_i\right)_\p \ne 0$. This can only happen if
    $\p_i\subset\p$, hence $\{\p_1,\dots,\p_n\}\subset \supp(Q)$, and
    these sets have the same minimal elements.

    Now we'll show that $\ass(Q)\subset\{\p_1,\dots,\p_n\}$. Let
    $\p\in\ass(Q)$, this means that
    \[
    R/\p \iso N \subset Q.
    \]
    Intersect $N$ with our finite filtration, and consider the
    smallest $i$ such that $N\cap Q_i\ne 0$. Write
    \[
    0=N\cap Q_{i-1}\to N\cap Q_i,
    \]
    and we must have that $R/\p \into (N\cap Q_i)/(N\cap Q_{i-1})$, and so
    \[
    R/\p \into Q_i/Q_{i-1} \iso R/\p_i.
    \]
    this can only happen if $\p_i\subset \p$, but assassins are
    maximal among annihilators, and so $\p_i=\p$. Thus
    \[
    \ass(Q)\subset\{\p_1,\dots,\p_n\}\subset \supp(Q).
    \]

    Finally, we will show that the minimal elements of $\supp(Q)$ are
    in $\ass(Q)$. Let $\p$ be a minimal element of
    $\supp(Q)$. Localize to obtain
    \[
    \supp(Q_\p) =\{\p R_\p\}.
    \]
    Since $\ass(Q_\p)\ne \emptyset$, and
    $\ass(Q_\p)\subset\supp(Q_\p)$, we have that $\p
    R_\p\in\ass(Q_\p)$. So by Proposition~\ref{P:assassinlocal},
    $\p\in\ass(Q)$.
  \end{proof}
\end{theorem}


\begin{corollary}[Assassins are finite]
  If $R$ is a Noetherian ring and $Q$ is a finitely generated
  $R$-module, then $\ass(Q)$ is finite.
  \begin{sketch}
    Our filtration above is finite.
  \end{sketch}
\end{corollary}


\begin{proposition}\label{P:subsetassassin}
  Let $R$ be a Noetherian ring and $L,M,N$ be finitely generated
  $R$-modules. Given a short exact sequence
  \[
  0 \to L \to M \to N \to 0,
  \]
  $\ass(L) \subset \ass(M) \subset \ass(L) \cup \ass(N)$.
  \begin{proof}
    Since $L\into M$, we see $\ass(L) \subset \ass(M)$.

    Suppose that $\p\in\ass(M)$, meaning $M$ contains a submodule $Q$
    isomorphic to $R/\p$.

    If $Q\cap L \ne \emptyset$, then there exists $x\in Q\cap L$ such
    that $\p = \ann(x)$, so $\p\in\ass(L)$.

    If $Q\cap L = 0$, $Q$ is isomorphic to a submodule of $M/L\iso N$,
    and so $\p\in\ass(N)$.
  \end{proof}
\end{proposition}








\begin{lemma}[Intersections and assassins]\label{L:intersectassassin}
  Consider a Noetherian ring $R$ and a finitely generated $R$-module
  $M$ with $R$-submodules $N_1,N_2\subset M$. If
  \[
  \ass(M/N_1) = \{\p\} \quad\text{and}\quad \ass(M/N_2) = \{\p\}
  \]
  then $\ass(M/(N_1\cap N_2)) = \{\p\}$.
  \begin{proof}
    The map
    \[
    M \to M/N_1\dsum M/N_2
    \]
    has kernel $N = N_1\cap N_2$. Hence
    \[
    M/(N_1\cap N_2)\into M/N_1\dsum M/N_2
    \]
    So
    \[
    \ass(M/(N_1\cap N_2)) \subset\ass(M/N_1\dsum M/N_2) = {\p}. 
    \]
    Thus, $\ass(M/(N_1\cap N_2))=\{\p\}$.
  \end{proof}
\end{lemma}


\end{document}



