\documentclass{ximera}

\input{../preamble.tex}

\author{Bart Snapp}

\title{Fundamental theorems}

\begin{document}
\begin{abstract}
  We review several fundamental theorems, all of which are special
  cases of primary decomposition. Sources and references:
  \cite{dE1995,hS2003}.
\end{abstract}
\maketitle

Let's review several fundamental theorems:

\begin{theorem}[Fundamental theorem of arithmetic]
  Every integer greater than $1$ can be factored uniquely (up to
  ordering) into primes.
\end{theorem}

Of course, you've now seen this in more generality.

%% \begin{theorem}
%%   Let $R$ be a euclidean domain\dots
%% \end{theorem}

\begin{theorem}[Fundamental theorem of finitely generated abelian groups]
  Every finitely generated abelian group $G$ is isomorphic to a
  direct sum of cyclic groups of prime-power order and infinite cyclic
  groups,
  \[
  G \iso \Z^n \dsum \directsum_{i=1}^{m} \Z_{p_i^{k_i}},
  \]
  where the $p_i$ are not necessarily distinct prime numbers.
\end{theorem}

%% \begin{theorem}[Structure theorem for finitely generated modules over a PID]
%%   Every finitely generated abelian group $G$ is isomorphic to a
%%   direct sum of cyclic groups of prime-power order and infinite cyclic
%%   groups,
%%   \[
%%   G \iso \Z^n \dsum \directsum_{i=1}^{m} \Z_{p_i^{k_i}},
%%   \]
%%   where the $p_i$ are not necessarily distinct prime numbers.
%% \end{theorem}

\begin{theorem}[Decomposing varieties]
  Let $V\subset\A^n$ be an affine algebraic variety over an
  algebraically closed field. We may express
  \[
  V = \bigcup_{i=1}^n U_i
  \]
  uniquely and minimally where each $U_i$ is an irreducible variety.
\end{theorem}


\begin{theorem}[Primary decomposition of ideals]
  Let $R$ be a Noetherian ring, and $I$ be an ideal of $R$. We may write
  \[
  I = \bigcap_{i=1}^n \q_i
  \]
  uniquely and minimally where each $\q$ is \dfn{primary}, meaning $\q\ne R$ and
  \[
  xy\in\q \quad \Rightarrow \quad x\in\q\text{ or }y^m\in\q.
  \]
\end{theorem}





\begin{definition}
  Let $R$ be a ring and $N$, $N_1$, and $N_2$ be proper submodules of
  some $R$-module $M$. The $R$-module $N$ is called \dfn{irreducible}
  if $N = N_1\cap N_2$ implies $N=N_1$ or $N=N_2$.
\end{definition}


\begin{exercise}
  Specialize the idea of an irreducible submodule to an irreducible
  ideal.
\end{exercise}

\begin{exercise}
  Specialize the idea of an irreducible submodule to an irreducible
  algebraic variety.
\end{exercise}


\begin{lemma}\label{L:irrdecomp}
  Let $R$ be a Noetherian ring and $M$ be a finitely generated
  $R$-module. The $R$-module $N\subsetneq M$ can be written as a finite
  intersection of irreducible submodules of $M$.
  \begin{sketch}
    Seek a contradiction by setting $\mathcal{S}$ to be the set of
    submodules of $M$ that cannot be expressed as a finite
    intersection of irreducible submodules. Use the fact that $M$ is
    Noetherian to conclude $\mathcal{S}$ must have a maximal element
    which cannot be irreducible. Find a contradiction.
  \end{sketch}
\end{lemma}

\begin{question}
What is the nature of the irreducible submodules described above?
\end{question}




\end{document}
