\documentclass{ximera}

\input{../preamble.tex}

\author{Bart Snapp}

\title{Primary decomposition}

\begin{document}
\begin{abstract}
  We prove the primary decomposition theorem. Sources and references:
  \cite{AM1969, sD2008, dE1995, mR1995}.
\end{abstract}
\maketitle


\begin{definition}
  Let $R$ be a ring and $N$, $N_1$, and $N_2$ be proper submodules of
  some $R$-module $M$. The $R$-module $N$ is called \dfn{irreducible}
  if $N = N_1\cap N_2$ implies $N=N_1$ or $N=N_2$.
\end{definition}


\begin{exercise}
  Specialize the idea of an irreducible submodule to an irreducible
  ideal.
\end{exercise}

\begin{exercise}
  Specialize the idea of an irreducible submodule to an irreducible
  algebraic varietiy.
\end{exercise}


\begin{lemma}\label{L:irrdecomp}
  Let $R$ be a Noetherian ring and $M$ be a finitely generated
  $R$-module. The $R$-module $N\subsetneq M$ can be written as a finite
  interesection of irreducible submodules of $M$.
  \begin{sketch}
    Seek a contradiction by setting $\mathcal{S}$ to be the set of
    submodules of $M$ that cannot be expressed as a finite
    intersection of irreducible submodules. Use the fact that $M$ is
    noetherian to conclude $\mathcal{S}$ must have a maximal element
    which cannot be irreducible. Find a contradiction.
  \end{sketch}
\end{lemma}

\begin{question}
What is the nature of the irreducible submodules described above?
\end{question}


\begin{definition}
  Let $R$ be a Noetherian ring. A nonzero finitely generated
  $R$-module $Q$ is called \dfn{coprimary} if for all $x\in R$, the
  homothety map (defined by multiplication by $x$)
  \[
  Q\lto^x Q
  \]
  is either injective or nilpotnent.
\end{definition}

\begin{exercise}
  Consider these questions and make wild conjectures:
  \begin{enumerate}
  \item Is the $\Z$-module $\Z_{18}$ is coprimary?
  \item Is the $\Z$-module $\Z_{9}$ is coprimary?
  \end{enumerate}
\end{exercise}


\begin{lemma}\label{L:irrcoprimary}
  Let $R$ be a Noetherian ring and $N\subset M$ be finitely generated
  $R$-modules. If $N$ is an irreducible submodule of $M$, then $Q=M/N$
  is coprimary.
  \begin{sketch}
    Construct an ascending chain of kernels of the homothety map
    defined via multiplication by $x\in R$. Use the fact that $R$ is
    Noetherian and $M$ is finitely generated.
    
    Supposing that $x q= 0$, we must show that $q= 0$ or that $x$ is
    nilpotent on $Q$.

    Let $a \in x^n Q \cap qR$. If $a\in qR$, argue $xa$ must be zero.
    Then suppose $a\in x^n Q$, so for $b\in Q$,
    \begin{align*}
      &\Rightarrow a = x^nb\\
      &\Rightarrow xa = x^{n+1}b\\
      &\Rightarrow x^n b = 0\\
      &\Rightarrow a=0.
    \end{align*}
    Use the fact that $0$ is an irreducible submodule of $Q$ to
    complete the proof.
  \end{sketch}
\end{lemma}


\begin{exercise}
  Specialize this lemma to ideals.
\end{exercise}



\begin{definition}
  Let $R$ be a ring and $Q$ be a $R$-module. The \dfn{assassins} of
  $Q$, denoted by $\ass(Q)$, is a set such that:
  \begin{itemize}
  \item $\ass(Q)\subset\spec(R)$.
  \item $Q$ contains a submodule isomorphic to $R/\p$.
  \end{itemize}
\end{definition}

\begin{exercise}
  Consider $\Z_6$ as a $\Z$-module. Find the assassins of $\Z_6$.
\end{exercise}

\begin{exercise}
  Let $k$ be a field, consider the $k[x]$-module $k[x]/(x^5)$. Find
  the assassins of $k[x]/(x^5)$.
\end{exercise}

\begin{exercise}
  Let $k$ be a field, consider the $k[x,y]$-module $k[x,y]/(xy)$. Find
  the assassins of $k[x,y]/(xy)$.
\end{exercise}


\begin{proposition}
  Let $Q$ be an $R$-module. If $\p\in\ass(Q)$, then $V(\p)\subset
  \supp(Q)$.
\end{proposition}

\begin{proposition}
  If $R$ is Noetherian, then $\ass(Q)$ contains a minimal element of
  $\supp(Q)$.
\end{proposition}




\begin{proposition}
  Let $R$ be a ring, $S$ be a multiplicacitvely closed subset of $R$,
  and $Q$ be an $R$-module. Then
  \[
  \p \in \ass_R(Q) \Leftrightarrow \ass_{S^{-1}}(S^{-1}(Q))
  \]
  whenever $S\cap\p = \emptyset$.
  \begin{sketch}
    $(\Rightarrow)$

    $(\Leftarrow)$
  \end{sketch}
\end{proposition}


\begin{lemma}
  If $R$ is a Noetherian ring and $Q$ is a finitely generated
  $R$-module, then $Q$ is coprimary if and only if it has exactly one
  assassin.
  \begin{sketch} 
    $(\Rightarrow)$ Suppose that $Q$ is a coprimary module. Use double
    containment to show any assassin is equal to $\rad{\ann(Q)}$.

    
    $(\Leftarrow)$ Seeking a contradiction, suppose there is a single
    assassin of $Q$, call it $\p$, but $Q$ is not coprimary . If $x\in
    \p$, but $x$ is not nilpotent on $Q$, localize at $x$. Contract
    the canonical map to find a second prime ideal in $\ass(Q)$.
  \end{sketch}
\end{lemma}


\begin{definition}
  A module is $Q$ is \dfn{$\boldsymbol{\p}$-coprimary} if $\ass(Q) = \{\p\}$.
\end{definition}

\begin{lemma}
  Let $Q$ be a finitely generated $\p$-coprimary $R$-module. Any
  nonzero submodule $T\subset Q$ is also $\p$-coprimary.
  \begin{sketch}
    %% $T$ is clearly coprimary, since the assassins are minimal, it must
    %% be $\p$-coprimary.
  \end{sketch}
\end{lemma}




\begin{definition}
  Given a ring $R$, an $R$-module $M$, and an $R$-submodule $N\subset
  M$, a \dfn{primary decomposition} of $N$ is
  \[
  N = \bigcap_{i=1}^n N_i
  \]
  where each $M/N_i$ is coprimary. 
\end{definition}


\begin{lemma}\label{L:intersectcoprimary}
  Given a ring $R$, an $R$-module $M$, and an $R$-submodules $N_1,N_2\subset
  M$, if $M/N_1$ and $M/N_2$ are $\p$-coprimary, then so is $M/(N_1\cap N_2)$.
\end{lemma}


\begin{theorem}[Primary decomposition]
  Let $R$ be a Noetherian ring. Given an $R$-module $M$ and an
  $R$-submodule $N\subset M$, there exist $R$-submodules $N_i\subset
  M$ such that
  \[
  N = \bigcap_{i=1}^n N_i
  \]
  where
  \begin{enumerate}
  \item $M/N_i$ is $\p_i$-coprimary for each $i$.
  \item If $i\ne j$, then $\p_i\ne\p_j$.
  \item $\ass(M/N) = \{\p_1,\dots,\p_n\}$.
  \end{enumerate}
  \begin{sketch}
    Use Lemmas~\ref{L:irrdecomp},~\ref{L:irrcoprimary},~\ref{L:intersectcoprimary},
    to show the first parts of the theorem.
    
    What remains is to show that $\ass(M/N) =
    \{\p_1,\dots,\p_n\}$. Use double containment.
    
    $(\subset)$ We will show that $\ass(M/N)
    \subset\{\p_1,\dots,\p_n\}$.
    \begin{align*}
      \frac{M}{N} = \frac{M}{N_1\cap\dots\cap N_n} &\lto^\eta \coprod_{i=1}^n M/N_i\\
        m &\mapsto (m + N_1,\dots, m+ N_n)
    \end{align*}
    Here $\ker(\eta) = N_1\cap\dots\cap N_n = N$, hence $\eta$ is an
    injection. Thus
    \[
    \ass(M/N) \subset \bigcup_{i=1}^n \ass(M/N_i) = \{\p_1,\dots,\p_n\}.
    \]


    $(\supset)$ We will show that $\{\p_1,\dots,\p_n\} \subset
    \ass(M/N)$. WLOG we'll show that $\p=\p_1\in\ass(M/N)$. Note
    \[
    \bigcap_{i=2}^n N_i/N \subset M/N
    \]
    and from before we have the injection
    \[
    \frac{M}{N_1\cap\dots\cap N_n} \lto^\eta \coprod_{i=1}^n M/N_i.
    \]
    This map can be restricted to
    \[
    \bigcap_{i=2}^n N_i/N \to \coprod_{i=1}^n M/N_i.
    \]
    however, the only nonzero part (and hence injective) part of this
    map is
    \[
    \bigcap_{i=2}^n N_i/N \into M/N_1.
    \]
    Since $\ass(M/N_1) = \{\p\}$, by Lemma~\ref{},
    $\ass\left(\bigcap_{i=2}^n N_i/N\right) = \{\p\}$, and so
    $\p\in\ass(M/N)$.
  \end{sketch}
\end{theorem}





\begin{theorem}
  Given a Noetherian ring $R$, and a finitely generated $R$-module $M$, we may embed
  \[
  M\into \coprod_{\p\in\ass(M)} Q(\p)
  \]
  where each $Q(\p)$ is a $\p$-coprimary module.
\end{theorem}













\end{document}



