\documentclass{ximera}

\input{../preamble.tex}

\author{Bart Snapp}

\title{Hensel's lemma}

\begin{document}
\begin{abstract}
  We state and prove Hensel's lemma. Sources and references:
  \cite{rA2006}.
\end{abstract}
\maketitle

\begin{theorem}[Hensel's lemma]\index{Hensel's lemma}
  Let $(A,\m,k)$ be a complete local ring and let $f$ be a monic
  polynomial of degree $d$ in $A[X]$. Consider the image of $f$ in a
  polynomial ring over the residue field, $\bar{f}\in k[X]$. If
  \begin{enumerate}
  \item $\bar{f} = \bar{g}\cdot \bar{h}$ in $k[X]$,
  \item with $(\bar{g},\bar{h}) = k[X]$,
  \end{enumerate}
  then $f = g\cdot h$ in $A[X]$, where $g= \bar{g}$ and $h=\bar{h}$
  are monic polynomials in $A[X]$.
  \begin{proof}
    We will work inductively, producing a sequence of polynomials,
    \[
    g_1,g_2,\dots \quad\text{and}\quad h_1,h_2,\dots
    \]
    in $A[X]$ where $\deg(g_i) = c$ and $\deg(h_i) = d-c$, such that
    \[
    f(X) - g_i(X) h_i(X) \in \m^i(X).
    \]

    
    If $i=1$, then by the hypothesis of the theorem, $\bar{f} =
    \bar{g}\cdot \bar{h}$ in $k[X]$. Thus
    \[
    f(X) - g_1(X) h_1(X) \in \m[X].
    \]

    Now assume our inductive construction holds up to $i=n$. Hence we have
    \[
    f(X) - g_n(X) h_n(X) = \sum_{i=0}^n a_i X^i\in \m^n[X].
    \]
    with $a_i\in \m^n$.  Since $(\bar{g}_n, \bar{h}_n) = k[X]$, there
    are polynomials $\bar{u}_i,\bar{v}_i\in k[X]$ such that
    \begin{align*}
      1  &= \bar{g}_n(X)\bar{u}_0(X) + \bar{h}_n(X)\bar{v}_0(X) \\
      X  &= \bar{g}_n(X)\bar{u}_1(X) + \bar{h}_n(X)\bar{v}_1(X) \\
      X^2  &= \bar{g}_n(X)\bar{u}_2(X) + \bar{h}_n(X)\bar{v}_2(X) \\
      &\ \vdots\\
      X^d  &= \bar{g}_n(X)\bar{u}_d(X) + \bar{h}_n(X)\bar{v}_d(X).
    \end{align*}
    Note that $\deg(\bar{u}_i) \le d-c$ and $\deg(\bar{v}_i) \le
    c$. If we lift $\bar{u}_i$ and $\bar{v}_i$, we have that
    \[
    X^i - g_n(X) u_i(X) - h_n(X) v_i(X) \in \m[X]. 
    \]
    Now set
    \begin{align*}
      g_{n+1}(X) &= g_n(X) + \sum_{i=0}^d a_i v_i(X)\\
      h_{n+1}(X) &= h_n(X) + \sum_{i=0}^d a_i u_i(X).
    \end{align*}
    Note that $\bar{g}_{n+1} = \bar{g}_n = g$ and $\bar{h}_{n+1} =
    \bar{h}_n = h$, and that the $\deg(\bar{g}_{n+1}) = c$ and
    $\deg(\bar{h}_{n+1}) = d-c$. We must show that
    \[
    f(X) - g_{n+1}(X) h_{n+1}(X) \in \m^{n+1}(X).
    \]
    Expanding out the element on the left,
    \[
    f(X) - \left(g_n(X) + \sum_{i=0}^d a_i v_i(X)\right)\left(h_n(X) + \sum_{i=0}^d a_i u_i(X)\right)
    \]
    yields
    \[
    \left(f(X) - g_n(X)h_n(X) - \sum_{i=0}^d a_i X^i\right) + \sum_{i=0}^d a_i(X^i - g_n(X) u_i(X) - h_n(X) v_i(X)) - T.
    \]
    By hypothesis, the left-most term is zero, by construction the
    middle term is in $\m^{n+1}[X]$, and the final term is a sum of
    elements of the form $a_ia_ju_i(X)v_i(X)\in\m^{n+1}[X]$.

    Now we will use that our ring is complete. The sequences
    \[
    g_1,g_2,\dots \quad\text{and}\quad h_1,h_2,\dots
    \]
    are such that $g_{n+1}-g_n\in \m^n[X]$ and $h_{n+1}-h_n\in
    \m^n[X]$. Thus the coefficients of each $X^i$ form a Cauchy
    sequence that converges. Let $g$ and $h$ be the polynomials with
    converged coefficients. Since $f$ is monic, the leading
    coefficients of $g$ and $h$ must be inverses of each other. Since
    the coefficients of these polynomials are not in $\m$, we may
    clear denominators to replace $g$ and $h$ with monic polynomials.
  \end{proof}
\end{theorem}


\section{Understanding \textit{p}-adic numbers}

We started this section with an algorithm for solving equations. This
algorithm is essentially Hensel's lemma. Thanks to Aziz Burak G\"ulen
for suggesting this approach.

\begin{corollary}[Hensel's lemma, \textit{p}-adic version]
  Let $f\in\Z[X]$ be a nonzero monic polynomial. Let $p$ be a prime in
  $\Z$ and $a_1$ be an integer such that
  \[
  f(a_1) \equiv 0 \pmod{p}\quad\text{with}\quad f'(a_1) \not\equiv 0\pmod{p}.
  \]
  Then there exists a sequence of integers $(a_1,a_2,\dots)$
  satisfying
  \[
  f(a_i) \equiv 0 \pmod{p^i}\quad\text{with}\quad a_{i+1} \equiv
  a_i\pmod{p^{i}}.
  \]
  The value of $a_i$ is determined uniquely modulo $p^i$.
  \begin{proof}
    Consider $f\in\Z[X]$. We will localize at a prime $p$ and complete to
    find
    \[
    f\in \Z[X] \subset \hat{\Z}_{(p)}[X].
    \]
    Now, suppose $\bar{f}(a_1) = (X-a_1)\cdot h(x)$ in
    $\Z/p\Z[X]$. Since $a_1$ is not a root of $f'$, $(X-a_1)$ and
    $h(X)$ have no common factors, so
    \[
    (X-a_1,h(X) = \Z/p\Z[X].
    \]
    by Hensel's lemma, $f = (X-a_1)h(X)$ in
    $\hat{\Z}_{(p)}[X]$. Iteratively repeating this process, we find a
    sequence of integers $(a_1,a_2,\dots)$ in $\hat{\Z}_{(p)}$
    corresponding to a solution in the $p$-adic integers.
  \end{proof}
\end{corollary}

This is the algorithm we used at the beginning of the section.


[Add a picture of $\hat{\Z}_{(3)}$.]

\section{Understanding the geometric case}

We will show how the prime spectrum changes upon completion.


[This section has yet to be written.]

For some interesting extra reading check out:
\begin{itemize}
\item \link[\textit{Visualizing the $p$-adic integers}, A.\ Cuoco, The American Mathematical Monthly
Vol.\ 98, No.\ 4 (Apr., 1991), pp.\ 355--364]{https://www.jstor.org/stable/2323809}. 

\item \link[\textit{On dividing a square into triangles}, P.\ Monsky, The American Mathematical Monthly
Vol.\ 77, No.\ 2 (Feb., 1970), pp.\ 161--164]{https://www.jstor.org/stable/2317329}. 
\end{itemize}





\end{document}
