\documentclass{ximera}


\input{../preamble.tex}

\author{Bart Snapp}

\title{Iteratively solving equations}

\begin{document}
\begin{abstract}
  We start by solving equations in using an iterative method.
\end{abstract}
\maketitle


Suppose you want to solve a polynomial equation $f(x)=0$ in $\Z$. You
have many options, but one option we will look at is based on the
following idea.


\begin{proposition}
  Let $x\in\Z$ and $p$ be a fixed prime in $\Z$. If $x$ is determined
  uniquely by its image in $\Z/p^n\Z$ for $n\ge 1$.
  \begin{sketch}
    Look for $n$ such that $p^{n-1}\le x < p^n$.
  \end{sketch}
\end{proposition}

One intuition for the proposition above is that for $n\gg 0$,
$\Z/p^n\Z$ looks a bit like $\Z$ when the elements of $\Z$ are near
zero.

Armed with this fact, suppose you want to solve a terrible equation in
$\Z$. As a toy example, we'll use
\[
x^2-36 x+323 = 0.
\]
Of course, this equation is easy to solve. We'll use a method that can
extend to a more general setting. Start by choosing a prime of $\Z$
where the equation has a solution in $\Z/p \Z$, we use $\Z/5\Z$, and
we'll call our first solution $x_1$. Write with me:
\begin{align*}
  x_1^2- 36x_1 + 323 &\equiv 0 &&\pmod{5}\\
  x_1^2 - x_1 + 3 &\equiv 0 &&\pmod{5}\\
  x_1 &\equiv 2&&\pmod{5}.
\end{align*}
In this case, $x_1 = 5y + 2$ for some $y\in \Z$. So now we will work
in $\Z/5^2\Z$. Write with me:
\begin{align*}
  x_1^2- 36x_1 + 323 &\equiv 0&&\pmod{5^2}\\
  15y+5 &\equiv 0 &&\pmod{5^2}\\
  y &\equiv -2 &&\pmod{5^2}
\end{align*}
Evaluating $x_1 = 5y+2$ at a representative of $y$, we obtain $x_2
\equiv -8 \pmod{5^2}$. This means for some new $y$, $x_2 = 5^2 y
-8$. Now we work in $\Z/5^3\Z$. Write with me:
\begin{align*}
  x_2^2- 36x_2 + 323 &\equiv 0&&\pmod{5^3}\\
  75y+50 &\equiv 0 &&\pmod{5^3}\\
  y &\equiv 1 &&\pmod{5^3}.
\end{align*}
Evaluating $x_2 = 5^2 y-8$ at a representative of $y$, we obtain $x_3
\equiv 17\pmod{5^3}$, so $x_3 = 5^3y+17$ for some new $y$. Now we work
in $\Z/5^4\Z$. Write with me:
\begin{align*}
  x_3^2- 36x_3 + 323 &\equiv 0 &&\pmod{5^4}\\
  375 &\equiv 0 &&\pmod{5^4}\\
  y &\equiv 0 &&\pmod{5^4}.
\end{align*}
Evaluating $x_3 = 5^3 y+17$ at a representative of $y$, we obtain $x_4
\equiv 17 \pmod{5^4}$.  We claim that continuing this process, we
would always obtain $17$ essence we have
\[
\begin{tikzcd}[row sep=0em,]
  \cdots \ar[r] & \Z/5^4\Z \ar[r] & \Z/5^3\Z \ar[r] & \Z/5^2\Z \ar[r] & \Z/5\Z \\
  & x_4\equiv 17 \ar[r,mapsto] & x_3\equiv 17 \ar[r,mapsto] & x_2\equiv -8 \ar[r,mapsto] &  x_1\equiv 2
\end{tikzcd}
\]
Let's check it:  
\[
(17)^2-36 (17)+323 = 0.
\]
OK, we've found a root.



\begin{question}
  How exactly does this algorithm work?
\end{question}

Let's try the algorithm now on an equation where we expect it to fail
in finding a solution. Say
\[
x^2 +  1 = 0
\]
over $\Z$. Again we'll choose the prime $p=5$. Write with me:
\begin{align*}
  x_1^2+1 &\equiv 0&&\pmod{5}\\
  x_1 &\equiv 2 &&\pmod{5}.
\end{align*}
In this case, $x_1 = 5y + 2$ for some $y\in \Z$. So now we will work
in $\Z/5^2\Z$. Write with me:
\begin{align*}
  x_1^2+1 &\equiv 0&&\pmod{5^2}\\
  20y+5 &\equiv 0 &&\pmod{5^2}\\
  y &\equiv 1 &&\pmod{5^2}
\end{align*}
Evaluating $x_1 = 5 y+2$ at a representative of $y$, we obtain $x_2
\equiv 7\pmod{5^2}$. So $x_2 = 5^2 y +7$. Now we work in
$\Z/5^3\Z$. Write with me:
\begin{align*}
  x_2^2+1 &\equiv 0&&\pmod{5^3}\\
  100y+50 &\equiv 0 &&\pmod{5^3}\\
  y &\equiv 2 &&\pmod{5^3}
\end{align*}
Evaluating $x_2 = 5^2 y+7$ at a representative of $y$, we obtain $x_3
\equiv 57\pmod{5^3}$. So $x_3 = 5^3 y +57$. Now we work in
$\Z/5^4\Z$. Write with me:
\begin{align*}
  x_3^2+1 &\equiv 0&&\pmod{5^4}\\
  500y+125 &\equiv 0 &&\pmod{5^4}\\
  y &\equiv 1 &&\pmod{5^4}
\end{align*}
Evaluating $x_3 = 5^3 y+57$ at a representative of $y$, we obtain $x_4
\equiv 182\pmod{5^4}$.  We claim that this process can continue. In
essence we have
\[
\begin{tikzcd}[row sep=0em,]
  \cdots \ar[r] & \Z/5^4\Z \ar[r] & \Z/5^3\Z \ar[r] & \Z/5^2\Z \ar[r] & \Z/5\Z \\
  & x_4\equiv 182 \ar[r,mapsto] & x_3\equiv 57 \ar[r,mapsto] & x_2\equiv 7 \ar[r,mapsto] &  x_1\equiv 2
\end{tikzcd}
\]
We now have some questions.  In particular, how exactly does the
algorithm work? What rings would it work for?  Let's see if we can
figure out what is going on.

\end{document}
