\documentclass{ximera}

\input{../preamble.tex}

\author{Bart Snapp}

\title{Topology and adic filtrations}

\begin{document}
\begin{abstract}
  We introduce topology to our study of adic filtrations. Sources and
  references: \cite{sD2008,jpS2000}.
\end{abstract}
\maketitle


\section{Topological rings and modules}

\begin{definition}\index{topological!ring}
  A ring $R$ is a \dfn{topological ring} if there exists some topology
  on $R$ such that the maps:
  \begin{align*}
    R\times R &\to R    & R\times R &\to R    \\
    (x,y) &\mapsto x+y  & (x,y) &\mapsto xy   
  \end{align*}
  are all continuous maps.
\end{definition}


\begin{definition}\index{topological!module}
  If $R$ is a topological ring, an $R$-module $M$ is a
  \dfn{topological module} if there exists some topology $M$ such that
  the maps:
  \begin{align*}
    M\times M &\to M    & R\times M &\to M \\
    (m,n) &\mapsto m+n  & (x,m) &\mapsto xm
  \end{align*}
  are both continuous maps.
\end{definition}


Now recall the open-set definition of a topology:

\begin{definition}
  A \dfn{topological space} is a set $X$ with a set of subsets called
  the \dfn{open sets} that satisfy the following properties:
  \begin{enumerate}
  \item The empty set and $X$ are open.
  \item Unions of arbitrary collections of open sets are
    open.
  \item Finite intersections of open sets are open.
  \end{enumerate}
\end{definition}

\begin{exercise}
  Prove the equivalence of the open set definition of a topology and
  the closed set definition of a topology.
\end{exercise}


While there is a more general presentation, we will stick to what we
are most interested in, adic topologies.


\begin{definition}
  Let $R$ be a ring, $I \subset R$ be an ideal, and $M$ an
  $R$-module. The \textbf{$\boldsymbol{I}$-adic topology} on $M$ is
  where $U \subset M$ is defined to be an open set if for each $x\in
  U$, there exists a positive integer $n$ such that
  \[
  x + I^n M \subset U.
  \]
\end{definition}



\begin{proposition}
  Let $R$ be a ring, $I\subset R$ be an ideal, and $M$ be a module
  equipped with the $I$-adic topology. If $N$ is a submodule of $M$,
  the closure $\bar{N}$ of $N$ is equal to
  \[
  \bigcap_{i=1}^\infty (N+I^n M).
  \]
  \begin{proof}
    $(\subset)$ If $x\in \bar{N}$, then it is in every
    neighborhood of $N$, and hence in the intersection.

    $(\supset)$ If $x\notin\bar{N}$, then there is $n$ such that
    $(x+I^nM)\cap N=\emptyset$ or $x\notin (N+I^nM)$.
  \end{proof}
\end{proposition}

\begin{corollary}
  Let $R$ be a ring, $I\subset R$ be an ideal, and $M$ be a module
  equipped with the $I$-adic topology. $M$ is Hausdorff if and only if
  \[
  \bigcap_{i=1}^\infty I^n M  = 0.
  \]
\end{corollary}

\begin{corollary}
  Let $(A,\m)$ be a local ring equipped with the $\m$-adic
  topology. In this case $A$ is Hausdorff.
\end{corollary}





\section{Metric spaces}


Recall the definition of a metric space:

\begin{definition}\index{pseudometric}\index{metric}
  A function of sets $d:M\times M\to [0,\infty)$ is called a \dfn{pseudometric} if:
    \begin{enumerate}
    \item For all $x,y\in M$, $d(x,y) = d(y,x)$.
    \item For all $x,y,z\in M$, $d(x,y) + d(y,z) \ge d(x,z)$.
    \end{enumerate}
    If in addition we have that for all $x,y \in M$, $d(x,y)=0$ if and
    only if $x = y$, then $d$ is called a \dfn{metric}.
\end{definition}

\begin{definition}
  A \dfn{metric space} is a set $M$ and a metric $d:M\times M \to
  [0,\infty)$.
\end{definition}

If $M$ has the $I$-adic topology, one may define a
pseudometric\index{pseudometric} on $M$ as follows: Fix any $c\in
(0,1)$. For any $x,y\in M$ define
\[
d(x,y) := c^n
\]
where $n$ is the integer such that $(x-y)\in I^nM-I^{n+1}M$, if no
such integer exists, then set $d(x,y) = 0$. If $M$ is
Hausdorff\index{Hausdorff}, then we have defined a metric.


\begin{corollary}
  The $\m$-adic topology induces a metric.
  \begin{proof}
    We know from Corollary~\ref{C:intmax}, that given a local ring $(A,\m)$,
    \[
    \bigcap_{n\ge 1} \m^n  = (0).
    \]
  \end{proof}
\end{corollary}

There are other cases where the $I$-adic topology induces a metric.

\begin{corollary}
  Let $R$ be a Noetherian ring and $I \subset \i J(R)$. If $M$ is a
  finitely generated module then
  \[
  \bigcap_{n=1}^\infty I^n M = 0.
  \]
  \begin{proof}
    By Artin-Rees, Theorem~\ref{L:ArtinRees},\index{Artin-Rees} there
    exists $k > 0$ such that for all $n$ we have
    \[
    N = N \cap I^{n + k}M = I^n(N \cap I^kM) = I^nN.
    \]
    Thus by \index{Nakayama's lemma}Nakayama's lemma,
    Theorem~\ref{NAK}, $N = 0$.
  \end{proof}
\end{corollary}


\begin{corollary}
  Let $A$ be a Noetherian domain. If $I$ is a proper ideal of $A$,
  then
  \[
  \bigcap_{n=1}^\infty I^n = 0.
  \]
  \begin{proof}
    Let $J = \bigcap I^n$.  Then we have $J = I J$.  By Nakayama's
    lemma, Theorem \ref{NAK}, there is a $x$ comaximal to $I$ such
    that $x\bigcap_{n=1}^\infty I^n=0$.  Since $A$ is a domain, we see
    $\bigcap_{n=1}^\infty I^n=0$.
  \end{proof}
\end{corollary}


\end{document}
