\documentclass{ximera}

\input{../preamble.tex}

\author{Bart Snapp}

\title{Affine space}


\begin{document}
\begin{abstract}
  We introduce affine space. Sources and references:
  \cite{mR1995, SKKT2002}.
\end{abstract}
\maketitle

The first topoloical space we study is known as \textit{affine space}.



\begin{definition}
  Let $k$ be a field and let $\vec{F} = \{F_{i}\}_{i\in I}$ be polynomials
  $k[\X]$, an \dfn{affine algebraic variety} is the set
  \[
  \V(\vec{F}) := \{\x: \text{$F_i(\x) = 0$ for all $i\in I$}\}.
  \]
\end{definition}


\begin{exercise}
  Give examples and nonexamples of algebraic varieties.
\end{exercise}





\begin{exercise}
  Let $\mathsf{M}$ be a $m\times m$ matrix with entries in some field
  $k$ and $\det(\mathsf{M}) \ne 0$. Is $\V( \mathsf{M} \vec{F}) =
  \V(\vec{F})$?
\end{exercise}


\begin{exercise}
  Show that if $I\subset J$, then $\V(J) \subset \V(I)$.
\end{exercise}

\begin{exercise}
  Let $F,G\in k[\X]$. Show that $\V(F) \cap\V(G) = \V(\{F,G\})$.
\end{exercise}


\begin{exercise}
  Let $F,G\in k[\X]$. Show that $\V(F) \cup\V(G) = \V(F\cdot G)$.
\end{exercise}





\begin{definition}
  Let $V$ be an algebraic variety defined by some polynomials in
  $k[\X]$. Define
  \[
  \I(V) = \{F\in k[\X]:F[\x] = 0 \text{ for } \x \in V\}.
  \]
\end{definition}

\begin{exercise}
  Given an algebraic variety defined by polynomials in $k[\X]$, show
  that $\I(V)$ is an ideal of $k[\X]$.
\end{exercise}


\begin{exercise}
  Let $\vec{F} = \{F_i\}_{i\in I}\subset k[\X]$. Show that every
  algebraic variety $\V(\vec{F})$ is the set of common zeros of
  finitely many polynomials in $k[\X]$.
\end{exercise}

\begin{exercise}
  Let $V$ be an algebraic variety defined by elements of $k[\X]$ and
  let $I$ be an ideal of $k[\X]$. Which of the following are true?
  \begin{enumerate}
  \item $\V(\I(V)) = V$
  \item $\I(\V(I)) = I$
  \end{enumerate}
\end{exercise}


This brings us to a question:


\begin{question}
  To what extent are varieties and ideals dual?
\end{question}

We will need to take some time to answer this question. Until then, we
press on.




\begin{proposition}
  The space $k^n$ where closed sets are defined to be the algebraic
  varieties is a topological space. This topology is called the
  \dfn{Zariski topology}, and when $k^n$ is equipped with this
  topology, we denote it $\A_k^n$, or simply $\A^n$.
  \begin{sketch}
    Check the conditions for a topology.
  \end{sketch}
\end{proposition}

\begin{exercise}
  Let $F\in k[X_1,\dots,X_n]$. Show that $F:\A^n\to \A^1$ is
  continuous. Hint: First show that the only closed subsets of $\A^1$
  are finite subsets and all of $\A^1$.
\end{exercise}


\begin{definition}
  Given an algebraic variety $\V(\vec{F})$, the \dfn{ring of regular
    functions} on $\V(\vec{F})$, is $k[\vec{X}]/(\vec{F})$.
\end{definition}

\begin{exercise}
  Show that the closed subsets of $\V(\vec{F})$ are of the form
  \[
  \{x\in \V(\vec{F}): G(x) = 0\text{ for all $G \in I$ an ideal of $k[\vec{X}]/(\vec{F})$}\}.
  \]
\end{exercise}


\begin{exercise}
  Given $G\in k[\vec{X}]/(\vec{F})$, show that $G: \V(\vec{F})\to
  \A^1$ is continuous with respect to the Zariski topology.
\end{exercise}



\begin{definition}
  Let $U\subseteq\A^m$ and $V\subseteq\A^n$ be affine algebraic
  varieties. A map $\vec{F} = (F_1,\dots,F_n)$
  \[
  \vec{F}: U\to V
  \]
  is a \dfn{morphism} of algebraic varieties if $F_1,\dots,F_n\in k[\X]/\I(U)$.
\end{definition}

\begin{exercise}
  Prove that every morphism of varieties is continuous with respect to
  the Zariski topology.
\end{exercise}

\begin{exercise}\label{E:affinemap}
   Let $U\subseteq\A^m$ and $V\subseteq\A^n$ be affine algebraic varieties. Show that a
   morphism $\vec{F}: U\to V$ induces a map of $k$-algebras
   \begin{align*}
     \vec{F}^*:  k[\X]/\I(V) &\to k[\X]/\I(U)\\
     G &\mapsto G\circ \vec{F}.
    \end{align*}
\end{exercise}


\begin{example}
	Let's look at a simple explicit example of an induced map of
        $k$-algebras given a map of varieties, as in
        Exercise~\ref{E:affinemap}. Consider the varieties
	\[
	U=\{(x,y)\in\mathbb C^2:y-x=0\} \quad\text{and}\quad
	V=\{(x,y)\in\mathbb C^2:y-2x=0\}
	\]
	and the map $\vec F:U\to V:(x,y)\mapsto(x,2y)$. Following the notation of Exercise~\ref{E:affinemap},
	\[
	\I(U)=(y-x) \quad\text{and}\quad
	\I(V)=(y-2x)
	\]
	so we wish to calculate
	\[
	\vec F^*:\mathbb C[x,y]/(y-2x)\to\mathbb C[x,y]/(y-x).
	\]
	Trying to write this map abstractly, we get
        \[
        G(x,y)\mapsto G(\vec F(x,y)=G(x,2y).
        \]
        We can explicitly map $G(x,y)=x^2+y^3x$ by
	\begin{equation*}
	\begin{tikzcd}
	x^2+y^3x \arrow[leftrightarrow]{r}{=}\arrow[mapsto]{d}{\vec F^*} & x^2+8x^4 \arrow[leftrightarrow]{r}{=}\arrow[mapsto]{d}{\vec F^*} & \tfrac14y^2+\tfrac12y^4 \arrow[mapsto]{d}{\vec F^*}\\
	x^2+8y^3x \arrow[leftrightarrow]{r}{=} & x^2+8x^4 \arrow[leftrightarrow]{r}{=} & y^2+8y^4
	\end{tikzcd}
	\end{equation*}
	so we see that our above interpretation of the map $\vec F^*$ really is well-defined.
\end{example}













\begin{proposition}%% From mR1995 5.7
  Show that an algebraic variety $V$ is irreducible if and only if
  $\I(V)$ is prime.
  %% \begin{sketch}
    
  %% \end{sketch}
\end{proposition}










This last propositon suggests that we might want to look into the set
of prime ideals of a ring.






















\end{document}
