\documentclass{ximera}

\input{../preamble.tex}

\author{Bart Snapp}

\title{The nullstellensatz}


\begin{document}
\begin{abstract}
  We prove Noether normalization, the weak nullstellensatz and the
  strong nullstellensatz. Sources and references:
  \cite{AM1969,dE1995,hM1986,mR1995}.
\end{abstract}
\maketitle


At this point we should believe there is a correspondence between:
affine algebraic varieties and ideals; and affine space and the prime
spectrum. Let's make this precise. We start by proving Noether
normalization. We will prove it now for infinite fields. Later we will
give a general proof.

\begin{theorem}[Noether normalization]
  Let $k$ be an infinite field and $A$ be a finitely generated
  $k$-algebra. There exist $\alpha_1,\dots,\alpha_m\in A$ such that
  \begin{enumerate}
  \item $\alpha_i$ are algebraically independent over $k$.
  \item $A$ is module-finite over $k[\boldsymbol{\alpha}]$.
  \end{enumerate}
  In other words, 
  \[
  k\subset k[\boldsymbol{\alpha}] \subset A,
  \]
  where $k[\boldsymbol{\alpha}]$ is a polynomial ring over $k$, and
  $A$ is a finitely generated as a $k[\boldsymbol{\alpha}]$-module.
  \begin{sketch}
    Proceed by induction on the number of generators
    $\zeta_1,\dots,\zeta_n$ of $A$ as a $k$-algebra. Consider
    \[
    k \subset B \subset k[\zeta] = A
    \]
    Decided what $B$ must be when $\zeta$ is algebraic over $k$, and
    when $\zeta$ is transendental over $k$.

    Now suppose that we know this theorem for $A$ generated by up to
    $n$ generators. We need to show this holds for $n$ generators.

    Consider the canonical map
    \[
    k[\X] \to k[\zeta]
    \]


    (More will be given later)
  \end{sketch}
\end{theorem}


Now we prove the so-called ``weak nullstellensatz.'' Again we will
insist that the field be infinite so that we can apply the version of
Noether normalization above. Later, when we prove Noether
normalization in generality, a more general version of the weak
nullstellensatz will also hold.

\begin{theorem}[Weak nullstellensatz]
  Let $k$ be a(n infinite) field, and $K$ be a finitely generated
  $k$-algebra that is also a field. In this case, $K$ is an algebraic
  extension of $k$.
  \end{theorem}




\begin{theorem}[Nullstellensatz]
  Let $k$ be an algebraically closed field with $I \subsetneq k[X_1,\dots,X_n]$.
  \begin{enumerate}
  \item $\V(I) \ne \emptyset$.
  \item $\I(\V(I)) = \rad{I}$.
  \end{enumerate}
\end{theorem}











\end{document}
