\documentclass{ximera}

\input{../preamble.tex}

\author{Bart Snapp}

\title{The prime spectrum}


\begin{document}
\begin{abstract}
  We introduce the prime spectrum. Sources and references:
  \cite{AM1969}.
\end{abstract}
\maketitle



\begin{definition}\index{nilpotent}
  An element $x$ of a ring is called \dfn{nilpotent} if there exists
  $n\in \N$ such that $x^n = 0$.
\end{definition}


\begin{exercise}
  Find the nilpotent elements of $\Z/n\Z$ where $n=pq$, $n=pq^2$ and
  $n = \prod_{i=1}^m p_i^{n_i}$; where $p$, $q$, $p_i$ are distinct
  primes of $\Z$.
\end{exercise}


\begin{exercise}
  If $u$ is a unit and $x$ is nilpotent, show that $u+x$ is a
  unit. Hint: Expand $(1+x/u)^{-1}$ as a power series.
\end{exercise}

\begin{definition}\index{radical0@$\rad{0}$}\index{nilradical}
  The set of nilpotent elements of a ring $R$ is called the
  \dfn{nilradical} of $R$. We will use $\rad{0}$ to denote this
  set. 
\end{definition}

\begin{exercise}
  Show that $\rad{0}$ is an ideal.
\end{exercise}


\begin{exercise}
  Let $R$ be a ring, and let $I\subset \rad{0}$. If $\bar{u}\in R/I$
  is a unit, prove that $u\in R$ is a unit.
\end{exercise}


We can play with the nilradical in \macaulay. Note, we need to insist
that $0$ is the zero of our ring:
\begin{macaulay2}
R = ZZ/101[x,y]/ideal(x^4,y^7)
radical ideal 0_R
\end{macaulay2}

We can generalize the idea of the nilradical as follows:

\begin{definition}\index{radical}
  The \dfn{radical} of an ideal $I$ is denoted by $\rad{I}$ and is
  defined to be the set
  \[
  \rad{I} = \{x\in R:x^n\in I\text{ for some }n\in\N\}.
  \]
\end{definition}

\begin{exercise}
  Show that $\rad{I}$ is an ideal.
\end{exercise}

\begin{exercise}
  If $I,J$ are ideals of $R$, and $\p$ is a prime ideal, the following
  hold:
  \begin{enumerate}
  \item $I \subset \rad{I}$.
  \item $\rad{I} =\rad{\rad{I}}$.
  \item $\rad{IJ}= \rad{I\cap J} = \rad{I} \cap \rad{J}$.
  \item $\rad{\p^n} = \p$.
  \end{enumerate}
\end{exercise}

We can play with the radicals in \macaulay:
\begin{macaulay2}
R = ZZ/101[x,y]
I = ideal(x^4,y^7)
J = ideal(y^3 - 3*y^2*x + 3*y*x^2 - x^3)
p = ideal(x^2+y^3+1)
isPrime p
rI=radical I
rJ=radical J
isSubset(I,rI)
rI == radical rI
radical(I*J)
radical intersect(I,J)
intersect(ideal rI,rJ)
radical (p^7)
\end{macaulay2}


\begin{proposition}
  Given a ring $R$, the radical of an ideal $I$ is equal to the
  intersection of all the prime ideals that contain $I$.
\end{proposition}

\begin{proof}
  $(\subset)$ Suppose that $x\in \rad{I}$. Then $x^n\in I$ and for
  each prime ideal containing $I$, $x^n\in \i p$. Since $\i p$ is
  prime, $x\in \i p$. Thus
  \[
  \rad{I} \subset \bigcap_{\i p \supset I}\i p.
  \]

  $(\supset)$ By the \index{Correspondence Theorem}Correspondence
  Theorem, the prime ideals of $R$ containing $I$ correspond
  bijectively to the ideals of $R/I$, hence we reduce to the case
  where $I =(0)$.

  Suppose that $x$ is not nilpotent. We'll show that
  \[
  x\notin \bigcap_{\p \supset (0)}\p.
  \]
  Consider the set of ideals of $R$:
  \[
  \mathcal{S} = \{\a:\text{$x^i \notin \a$ for $i> 0$}\}
  \]
  Note that $(0)\in \mathcal{S}$ and that $\mathcal{S}$ may be ordered
  by inclusion. Now let $\mathcal{C}$ be any chain of ideals in
  $\mathcal{S}$. This chain has an upper bound in $\mathcal{S}$,
  namely the ideal:
  \[
  \bigcup_{\a\in\mathcal{C}} \a
  \]
  Hence by Zorn's Lemma,\index{Zorn's Lemma} $\mathcal{S}$ has a
  maximal element, call it $\p$. We claim that $\p$ is prime. Suppose
  that $a,b\notin\p$. Hence $(a) + \p$ and $(b)+\p$ are ideals not
  contained in $\mathcal{S}$. Thus for some $m,n\in\N$:
  \[
  x^m \in (a) + \p \qquad\text{and}\qquad x^n \in (b) + \p 
  \]
  Moreover,
  \[
  x^{m+n}\in (ab) + \p
  \]
  and so we see that $ab\notin\p$. Hence $\p$ is prime and
  $x\notin\p$. Thus $x$ is not in the intersection of the prime ideals
  of $R$.
\end{proof}


\begin{definition}
  Let $R$ be a ring, the \dfn{prime spectrum} of $R$ is the set of all
  prime ideals of $R$. We denote the prime spectrum by $\spec(R)$.
\end{definition}

\begin{definition}
  Let $R$ be a ring, then $V(E)\subset \spec(R)$ is the set of all
  prime ideals that contain $E$.
\end{definition}

\begin{exercise}
  Show that if $I\subset J$, then $V(J) \subset V(I)$.
\end{exercise}

\begin{proposition}
  The set $\spec(R)$ is a topological space with closed sets
  \[
  \{ V(E):E\subset R\}.
  \]
  This topology is called the \dfn{Zariski topology}.
  \begin{sketch}
    Check the conditions for a topology.
  \end{sketch}
\end{proposition}

\begin{exercise}%% From AM1969, Ch 1 (16)
  Draw pictures of:
  \begin{enumerate}
  \item $\spec(\Z)$
  \item $\spec(\R)$
  \item $\spec(\C[x])$
  \item $\spec(\R[x])$
  \item $\spec(\Z[x])$
  \end{enumerate}
\end{exercise}


\begin{exercise}
  For each $f\in R$, set
  \[
  U_f := \spec(R) - V(f).
  \]
  Show that the sets $U_f$ form a basis for $\spec(R)$, that is show:
  \begin{enumerate}
    \item For each $\p\in\spec(R)$, there is some $f\in R$ such that
      $\p\in U_f$.
    \item If $\p\in U_f \cap U_g$, then there is $h\in R$ such that
      $U_h\subseteq U_f\cap U_g$, and $\p\in U_h$.
  \end{enumerate}
\end{exercise}


\begin{exercise}%% From AM1969, Ch 1 (17)
  Let $U_f$ be defined as above. Prove that:
  \begin{enumerate}
  \item $U_f \cap U_g = U_{fg}$
  \item $U_f = \emptyset$ if and only if $f$ is nilpotent.
  \item $U_f = \spec(R)$ if and only if $f$ is a unit.
  \item $U_f = U_g$ if and only if $\rad{f} = \rad{g}$.
  \item $\spec(R)$ is quasi-compact (every open covering of $\spec(R)$
    has a finite subcovering, but $\spec(R)$ might not be Hausdorff).
  \end{enumerate}
\end{exercise}


\begin{exercise}%% From AM1969, Ch 1 (21)
  Let $\phi:R \to S$ be a homomorphism of rings. For $P\in\spec(S)$,
  note that $\phi^{-1}(P) \in \spec(R)$. Hence there is an induced map
  \[
  \phi^*:\spec(S) \to \spec(R). 
  \]
  Prove that:
  \begin{enumerate}
  \item $\phi^*$ is continuous.
  \item If $\a$ is an ideal of $R$, then ${\phi^*}^{-1}(V(\a)) =
    V(\phi(\a) S)$.
  \item If $\b$ is an ideal of $S$, then $\bar{\phi^*(V(\b))} =
    V(\phi^{-1}(\b)R)$.
  \item If $\phi$ is surjective, then $\phi^*$ is a homeomorphism from
    $\spec(S)$ to $V(\ker(\phi))$.
  \item $\spec(R)$ and $\spec(R/\rad{0})$ are homeomorphic.
  \item If $\phi$ is injective, then $\phi^*(\spec(S))$ is dense in
    $\spec(R)$.
  \item $\phi^*(\spec(S))$ is dense in $\spec(R)$ if and only if
    $\ker(\phi)\subseteq \rad{0}$.
  \item If
    \[
    R\lto^\phi S \lto^\psi T,
    \]
    then
    \[
    \begin{tikzcd}[column sep=2em]
      \spec(T) \ar[r, "\psi^*"] \ar[rr, "(\phi \circ \psi)^*"', bend right=20]
      & \spec(S) \ar[r, "\phi^*"]
      & \spec(R).
    \end{tikzcd}
    \]
  \item Let $R$ be an integral domain with a single prime ideal $\p$.
    Define
    \begin{align*}
      \phi:R &\to R/\p \times \ff(R)\\
      x &\mapsto (\bar{x}, x).
    \end{align*}
    Show that $\phi^*$ is bijective, but not a homeomorphism from
    $\spec(R/\p \times \ff(R))$ to $\spec(R)$.
  \end{enumerate}
\end{exercise}


\begin{exercise}%% From AM1969, Ch 1 (19)
  Show that $V(I)$ is irreducible if and only if $\rad{I}$ is a prime
  ideal.
\end{exercise}



I think we have another pressing question:

\begin{question}
  What is the connection between affine space and the prime spectrum?
\end{question}


\end{document}
