\documentclass{ximera}

\input{../preamble.tex}

\author{Bart Snapp}

\title{The support of a module}


\begin{document}
\begin{abstract}
  We introduce the support of a module. Sources and references:
  \cite{AM1969, mR1995}.
\end{abstract}
\maketitle

Let's take an aside, and think about the set of all continuous
functions mapping $\R$ to itself.

\begin{definition}
  Let $X$ and $Y$ be topological spaces. The set $\mathcal{C}(X,Y)$
  will be the continuous functions from $X$ to $Y$.
\end{definition}

\begin{exercise}
  Show that $\mathcal{C}(\R,\R)$ is a ring.
\end{exercise}

If we want to understand $\mathcal{C}(\R,\R)$ locally, we want to view
functions in a neighborhood of a specific point, say $x=3$. To do
this, let's create equivalence classes of continuous functions if
those functions agree on local neighborhoods of $x=3$.

\begin{definition}
  Let $p\in X$, we'll call this set of equivalence classes of
  continuous functions that agree on a neighborhood of $p$ the
  \textit{germs of continuous functions at $p$}, and denote this by
  $\mathcal{C}(X,Y)_p$.
\end{definition}


\begin{exercise}
  Show that $\mathcal{C}(\R,\R)_{x=3}$ is a local ring whose maximal
  ideal consists of the germs that are zero when evaluated at $x=3$.
\end{exercise}


\begin{definition}
  The \dfn{support} of a function $f:A\to B$ is the set of points in $A$
  where $f$ is nonzero. 
\end{definition}

The continuous functions (morphisms) on an affine algebraic variety
$V$ are elements of $k[\X]/\I(V)$, a $k[\X]$-module. While the points
of affine space correspond to maximal ideals, it is more convenient to
think about irreducible varieties. Hence we arrive a the definition of
the support of a module.


\begin{definition}
  Let $M$ be an $R$-module. The \dfn{support} of $M$ is a subset of $\spec(R)$ where
  \[
  \supp(M) = \{\p\in\spec(R):M_\p \ne 0\}.
  \]
\end{definition}



\begin{exercise}
  Let $M$ be an $R$-module. If $\p\in\supp(M)$, then
  $V(\p)\subseteq\supp(M)$.
\end{exercise}



\begin{proposition}%% AM and MR
  Let $R$ be a ring and $L,M,N$ be $R$-modules.
  \begin{enumerate}
  \item If $M = mR$, then $\supp(M) = V(\ann(m))$.
  \item If $M = \directsum_{i\in \mathcal I} M_i$, then $\supp(M) = \union_{i\in \mathcal I} \supp(M_i)$.
  \item Given a short exact sequence
    \[
    0 \to L \to M \to N \to 0,
    \]
    $\supp(M) = \supp(L) \cup \supp(N)$.
  \end{enumerate}
\end{proposition}

\begin{exercise}
  If $M$ is a finitely generated $R$-module, then $\supp(M) = V(\ann(M))$.  This
  shows that the support of any finitely generated $R$-module is a
  closed subset of $\spec(R)$.

  Now consider the $\Z$-module $M = \coprod_{\p\in U} \Z/\p\Z$ where $U$
  is a subset of $\spec(\Z)$ that is not closed. Explain why the
  support of $M$ is not a closed subset of $\spec(\Z)$.
\end{exercise}



%% some discussion sheaves might be nice here...


%% \begin{exercise}
%%   Given a morphism $\Phi:\supp(M) \to \supp(N)$, can you induce
%%   \[
%%   \Phi^* : N\to M?
%%   \]
%% \end{exercise}














\end{document}
